### 2024-11-29 paper/paper.qmd
updated tests and fixed data cleaning script

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Kevin Roe
thanks: "Code and data are available at: [https://github.com/Kanghyunroe/traffic_collisions/tree/main](https://github.com/Kanghyunroe/traffic_collisions/tree/main)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-29 scripts/00-simulate_data.R

@@ -9,6 +9,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 set.seed(690)
 
 #### Simulate data ####
@@ -35,44 +36,46 @@ simulated_data <- tibble(
   # Randomly Assign Months
   month = sample(list_of_months, size = 1000, replace = TRUE),
   
-  
   # Randomly Assign Day of the Week
   date_of_week = sample(list_of_dates, size = 1000, replace = TRUE),
   
   # Randomly Assign Years
   year = sample(2014:2024, size = 1000, replace = TRUE), 
   
+  # Randomly Assign Hours
+  hour = sample(0:24, size = 1000, replace = TRUE),
+  
   # Randomly Assign Police Divisions
-  date_of_week = sample(list_of_divisions, size = 1000, replace = TRUE),
+  police_division = sample(list_of_divisions, size = 1000, replace = TRUE),
   
   # Randomly Fatalities Variable
-  fatalities <- sample(c(0, 1), size = 1000, replace = TRUE),
+  fatalities = sample(0:1, size = 1000, replace = TRUE),
   
   # Randomly Injury Collision Variable
-  injury_collision <- sample(c(0, 1), size = 1000, replace = TRUE),
+  injury_collision = sample(0:1, size = 1000, replace = TRUE),
   
   # Randomly fail_to_remain_collision Variable
-  fail_to_remain_collision <- sample(c(0, 1), size = 1000, replace = TRUE),
+  fail_to_remain_collision = sample(0:1, size = 1000, replace = TRUE),
   
   # Randomly property damage collision Variable
-  property_damage_collision <- sample(c(0, 1), size = 1000, replace = TRUE),
+  property_damage_collision = sample(0:1, size = 1000, replace = TRUE),
   
   # Randomly automobile Variable
-  automobile <- sample(c(0, 1), size = 1000, replace = TRUE),
+  automobile = sample(0:1, size = 1000, replace = TRUE),
   
   # Randomly motorcycle Variable
-  motorcycle <- sample(c(0, 1), size = 1000, replace = TRUE),
+  motorcycle = sample(0:1, size = 1000, replace = TRUE),
   
   # Randomly passenger Variable
-  passenger <- sample(c(0, 1), size = 1000, replace = TRUE),
+  passenger = sample(0:1, size = 1000, replace = TRUE),
 
   # Randomly bicycle Variable
[…]

### 2024-11-29 scripts/01-test_simulated_data.R

@@ -12,29 +12,31 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(here)
+library(arrow)
 
-test_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+test_data <- read_parquet(file = here("data/00-simulated_data/simulated_data.parquet"))
 
 #### Test data ####
 
 # Check column class 
 class(test_data$month) == "character"
 class(test_data$date_of_week) == "character"
-class(test_data$year) == "number"
-class(test_data$hour) == "number"
+class(test_data$year) == "numeric"
+class(test_data$hour) == "numeric"
 class(test_data$police_division) == "character"
 class(test_data$fatalities) == "number"
-class(test_data$fail_to_remain_collision) == "number"
-class(test_data$property_damage_collision) == "number"
-class(test_data$automobile) == "number"
-class(test_data$motorcycle) == "number"
-class(test_data$passenger) == "number"
-class(test_data$bicycle) == "number"
-class(test_data$pedestrian) == "number"
+class(test_data$fail_to_remain_collision) == "numeric"
+class(test_data$property_damage_collision) == "numeric"
+class(test_data$automobile) == "numeric"
+class(test_data$motorcycle) == "numeric"
+class(test_data$passenger) == "numeric"
+class(test_data$bicycle) == "numeric"
+class(test_data$pedestrian) == "numeric"
 
 # Test that the Dataset has 15 columns
 test_that("dataset has 15 columns", {
-  expect_equal(ncol(test_data), 16)
+  expect_equal(ncol(test_data), 15)
 })
 
 # Test that 'month' column contains valid months 
@@ -57,7 +59,7 @@ test_that("'month' contains valid month names or NaN", {
 })
 
 # Test that 'hours' column contains valid hours
-valid_hours <- 1:24
+valid_hours <- 0:24
 
 test_that("'hour' contains valid hours (1-24)", {
   expect_true(all(is.na(test_data$hour) | test_data$hour %in% valid_hours))

### 2024-11-29 scripts/03-clean_data.R

@@ -42,49 +42,54 @@ cleaned_data <-
          "pedestrian" = "PEDESTRIAN", 
   ) |>
   
-  # Update Fatalities and Clean Other Variables
-  mutate(fatalities = ifelse(is.na(fatalities), 0, fatalities)) |>
+  # Update Fatalities to be Binary Variable 
+  mutate(fatalities = ifelse(is.na(fatalities), 0, 1)) |>
   
   mutate(
     injury_collision = case_when(
       injury_collision == "YES" ~ 1,
       injury_collision == "NO" ~ 0,
-      injury_collision == "N/R" ~ NA_real_
+      injury_collision == "None" ~ 0
     ),
     fail_to_remain_collision = case_when(
       fail_to_remain_collision == "YES" ~ 1,
       fail_to_remain_collision == "NO" ~ 0,
-      fail_to_remain_collision == "N/R" ~ NA_real_
+      fail_to_remain_collision == "None" ~ 0
     ),
     property_damage_collision = case_when(
       property_damage_collision == "YES" ~ 1,
       property_damage_collision == "NO" ~ 0,
-      property_damage_collision == "N/R" ~ NA_real_
+      property_damage_collision == "None" ~ 0
     ),
     automobile = case_when(
       automobile == "YES" ~ 1,
       automobile == "NO" ~ 0,
-      automobile == "N/R" ~ NA_real_
+      automobile == "N/R" ~ NA_real_,
+      automobile == "None" ~ 0
     ),
     motorcycle = case_when(
       motorcycle == "YES" ~ 1,
       motorcycle == "NO" ~ 0,
-      motorcycle == "N/R" ~ NA_real_
+      motorcycle == "N/R" ~ NA_real_,
+      motorcycle == "None" ~ 0
     ),
     passenger = case_when(
       passenger == "YES" ~ 1,
       passenger == "NO" ~ 0,
-      passenger == "N/R" ~ NA_real_
+      passenger == "N/R" ~ NA_real_,
+      passenger == "None" ~ 0
     ),
     pedestrian = case_when(
       pedestrian == "YES" ~ 1,
       pedestrian == "NO" ~ 0,
-      pedestrian == "N/R" ~ NA_real_
+      pedestrian == "N/R" ~ NA_real_,
+      pedestrian == "None" ~ 0
     ), 
     bicycle = case_when(
       bicycle == "YES" ~ 1,
[…]

### 2024-11-29 scripts/04-test_analysis_data.R

@@ -12,8 +12,9 @@
 library(tidyverse)
 library(testthat)
 library(arrow)
+library(here)
 
-test_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
+test_data <- read_parquet(file = here("data/02-analysis_data/analysis_data.parquet"))
 
 
 #### Test data ####
@@ -21,21 +22,21 @@ test_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 # Check column class 
 class(test_data$month) == "character"
 class(test_data$date_of_week) == "character"
-class(test_data$year) == "number"
-class(test_data$hour) == "number"
+class(test_data$year) == "numeric"
+class(test_data$hour) == "numeric"
 class(test_data$police_division) == "character"
 class(test_data$fatalities) == "number"
-class(test_data$fail_to_remain_collision) == "number"
-class(test_data$property_damage_collision) == "number"
-class(test_data$automobile) == "number"
-class(test_data$motorcycle) == "number"
-class(test_data$passenger) == "number"
-class(test_data$bicycle) == "number"
-class(test_data$pedestrian) == "number"
+class(test_data$fail_to_remain_collision) == "numeric"
+class(test_data$property_damage_collision) == "numeric"
+class(test_data$automobile) == "numeric"
+class(test_data$motorcycle) == "numeric"
+class(test_data$passenger) == "numeric"
+class(test_data$bicycle) == "numeric"
+class(test_data$pedestrian) == "numeric"
 
 # Test that the Dataset has 15 columns
 test_that("dataset has 15 columns", {
-  expect_equal(ncol(test_data), 16)
+  expect_equal(ncol(test_data), 15)
 })
 
 # Test that 'month' column contains valid months 
@@ -58,7 +59,7 @@ test_that("'month' contains valid month names or NaN", {
 })
 
 # Test that 'hours' column contains valid hours
-valid_hours <- 1:24
+valid_hours <- 0:24
 
 test_that("'hour' contains valid hours (1-24)", {
   expect_true(all(is.na(test_data$hour) | test_data$hour %in% valid_hours))

### 2024-11-29 scripts/05-model_data.R
added interaction terms

@@ -22,25 +22,22 @@ analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 analysis_data$month <- factor(analysis_data$month)
 analysis_data$police_division <- factor(analysis_data$police_division)
 
-set.seed(420)
-
-# get the reduced dataset of only 1000 randomly selected data entries
+# Get the reduced dataset of only 1000 randomly selected data entries
 motor_fatality_reduced_data <- slice_sample(analysis_data, n = 2000)
+
+set.seed(420)
 
 motor_fatality_prediction_model <-
   stan_glm(
-    fatalities ~ hour + date_of_week + month + 
-      police_division + 
-      property_damage_collision + automobile + motorcycle + passenger + 
-      bicycle + pedestrian,
+    fatalities ~ hour + injury_collision + fail_to_remain_collision + property_damage_collision + automobile + 
+      motorcycle + passenger + bicycle + pedestrian + police_division + hour*automobile + hour*pedestrian + hour* motorcycle + hour*passenger + hour*bicycle,
     data = motor_fatality_reduced_data,
     family = binomial(link = "probit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = 
       normal(location = 0, scale = 2.5, autoscale = TRUE),
     seed = 420,
-    iter = 4000,
-    warmup = 2000,
+    iter = 2000,
     init = "0"
   )
 

### 2024-11-29 scripts/05-model_data.R
added base model for comparison

@@ -20,6 +20,7 @@ analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
 # Convert variables to factors
 analysis_data$month <- factor(analysis_data$month)
+analysis_data$year <- factor(analysis_data$year)
 analysis_data$police_division <- factor(analysis_data$police_division)
 
 # Get the reduced dataset of only 1000 randomly selected data entries
@@ -27,10 +28,24 @@ motor_fatality_reduced_data <- slice_sample(analysis_data, n = 2000)
 
 set.seed(420)
 
+motor_fatality_base_model<-
+  stan_glm(
+    fatalities ~ hour + injury_collision + fail_to_remain_collision + property_damage_collision + automobile + 
+      motorcycle + passenger + bicycle + pedestrian,
+    data = motor_fatality_reduced_data,
+    family = binomial(link = "probit"),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = 
+      normal(location = 0, scale = 2.5, autoscale = TRUE),
+    seed = 420,
+    iter = 2000,
+    init = "0"
+  )
+
 motor_fatality_prediction_model <-
   stan_glm(
     fatalities ~ hour + injury_collision + fail_to_remain_collision + property_damage_collision + automobile + 
-      motorcycle + passenger + bicycle + pedestrian + police_division + hour*automobile + hour*pedestrian + hour* motorcycle + hour*passenger + hour*bicycle,
+      motorcycle + passenger + bicycle + pedestrian + police_division + year + hour*automobile + hour*pedestrian + hour* motorcycle + hour*passenger + hour*bicycle,
     data = motor_fatality_reduced_data,
     family = binomial(link = "probit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -43,14 +58,26 @@ motor_fatality_prediction_model <-
 
 
 #### Save model ####
+summary(motor_fatality_base_model)
 summary(motor_fatality_prediction_model)
+
+saveRDS(
+  motor_fatality_base_model,
+  file = "models/motor_fatality_base_model.rds"
+)
+
 saveRDS(
   motor_fatality_prediction_model,
   file = "models/motor_fatality_prediction_model.rds"
 )
 
-
 #### Posterior Model Checks ####
+posterior_predict(motor_fatality_base_model)
+pp_check(motor_fatality_base_model)
+
[…]

### 2024-11-29 scripts/05-model_data.R
improved code readability

@@ -28,6 +28,7 @@ motor_fatality_reduced_data <- slice_sample(analysis_data, n = 2000)
 
 set.seed(420)
 
+# Base Model
 motor_fatality_base_model<-
   stan_glm(
     fatalities ~ hour + injury_collision + fail_to_remain_collision + property_damage_collision + automobile + 
@@ -41,6 +42,8 @@ motor_fatality_base_model<-
     iter = 2000,
     init = "0"
   )
+
+# Extension Model with Interaction and Fixed Effects
 
 motor_fatality_prediction_model <-
   stan_glm(
@@ -57,7 +60,7 @@ motor_fatality_prediction_model <-
   )
 
 
-#### Save model ####
+#### Save models ####
 summary(motor_fatality_base_model)
 summary(motor_fatality_prediction_model)
 
@@ -72,12 +75,15 @@ saveRDS(
 )
 
 #### Posterior Model Checks ####
+
+# Base Model Checks
 posterior_predict(motor_fatality_base_model)
 pp_check(motor_fatality_base_model)
 
 base_posterior_summary <- describe_posterior(motor_fatality_base_model)
 print(base_posterior_summary)
 
+# Extension Model Checks
 posterior_predict(motor_fatality_prediction_model)
 pp_check(motor_fatality_prediction_model)
 

### 2024-12-01 scripts/05-model_data.R
updated model from probit to logit regression

@@ -34,7 +34,7 @@ motor_fatality_base_model<-
     fatalities ~ hour + injury_collision + fail_to_remain_collision + property_damage_collision + automobile + 
       motorcycle + passenger + bicycle + pedestrian,
     data = motor_fatality_reduced_data,
-    family = binomial(link = "probit"),
+    family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = 
       normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -50,7 +50,7 @@ motor_fatality_prediction_model <-
     fatalities ~ hour + injury_collision + fail_to_remain_collision + property_damage_collision + automobile + 
       motorcycle + passenger + bicycle + pedestrian + police_division + year + hour*automobile + hour*pedestrian + hour* motorcycle + hour*passenger + hour*bicycle,
     data = motor_fatality_reduced_data,
-    family = binomial(link = "probit"),
+    family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = 
       normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-12-01 paper/paper.qmd
added data outline and introduction draft

@@
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 library(tidyverse)
+library(here)
+library(arrow)
+library(ggplot2)
+library(modelsummary)
+library(rstanarm)
+library(knitr)
+library(kableExtra)
+
+# Read in Analysis Data
+analysis_data <- read_parquet(file = here("data/02-analysis_data/analysis_data.parquet"), show_col_types = FALSE)
+
+# Read in Model Data 
+prediction_model <- readRDS(file = here(
+  "models/motor_fatality_prediction_model.rds"))
 ```
@@
 
-Overview paragraph
+Automobile fatalities are a leading cause of death worldwide, posing a public health and safety concern for urban cities.
+For example, in 2024, thirty people have died on Toronto's roadways so far, which is a 20% increase from last year (insert citation, CBC).
+However, new headlines, such as CBCs, highlight that Motor Vehicle Collisions (MVC) tend to either be generalized as a summary statistic or typically fatal events are over analyzed to the extent that environmental factors surrounding the crash are ignored (CLARIFY).
+Moreover, general environmental factors such as the time of the crash or if a bicycle was involved in the accident are important to understand what increases the likelihood of a fatality occurring in an MVC.
+The use of statistical modeling on increasingly available vehicle collision data presents an opportunity to develop a nuanced understanding of what factors increases the likelihood for a fatality to occur in an accident.
+This paper uses the Toronto Police Service's  Annual Statistical Report from Open Data Toronto to analyze what factors are most responsible in predicting if a fatality occurs in a car crash.
 
-Estimand paragraph
+The estimand of interest is the log-adjusted probability of a fatality occuring in a car crash.
+Specifically, we aim to quanitfy how specific environmental factors increase or decrease the likelihood of a fatality.
+By applying inferential analysis through Bayesian linear models, we assess not only the magnitude of these effects, but their underlying uncertainties (EDIT).
 
-Results paragraph
+what was found (FINISH PARAGRAPH)
 
-Why it matters paragraph
+The paper is not only important from a public health perspective, but the paper also has policy development implications.
+Road safety and reducing fatal MVCs are a critical agenda item of any municipal government.
+The paper informs what factors increase the likelihood of death in an MVC, which informs policymakers' focus for relevant policy design.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The remainder of this paper is structured as follows: @sec-data describes the dataset and methodology and @sec-model exhibits the use of inferential models.
+@Sec-results presents the results of the analysis, detailing the observed relationships between likelihood of death and various circumstantial factors.
+@sec-discussion discusses the broader implications and limitations of our findings.@sec-appendix presents a detailed idealized methodology to improve data collection, and additional model summary and diagnostic information.
 
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+This dataset, "Police Annual Statistical Report - Traffic Collisions", was published and refreshed on October 21st, 2024, by the Toronto Police Service [insert citation].
+The Toronto Police Service publishes various datasets on public safety and crime to inform the public about safety issues [@annual_statistics_report].
+Data on traffic collisions is included in the Toronto Police Service's Annual Statistical Report, which also covers reported crimes, search of persons, firearms, and the Police Service's budget [@annual_statistics_report].
+The data is collected using historical Motor Vehicle Collisions and classifies them into the following categories:
+* Property Damage (PD) Collisions
+* Fail to Remain (FTR) Collisions, or commonly known as hit-and-run accidents
+* Injury Collisions
+* Fatalities
 
-Overview text
+Following the Municipal Freedom of Information and Protection of Privacy Act, the Toronto Police Service ensures to protect the privacy of individuals involved in the reported crimes when publishing the data.
+The dataset is updated annually, is open data, and can be used if an attribution statement @sec-appendix-attribution and is properly cited [@tphlicense].
+
+There is an alternative dataset from the Toronto Police Service called "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" (CITE).
+Unlike the alternative dataset, this paper's dataset focuses on all collisions, instead of only focusing on ones where someone was either killed or seriously injured.
+While the alternative dataset has more explanatory variables simply because more data is collected when someone dies or is seriously injured, this paper's aims to generalize if a fatality is more likely to occur based on the general circumstances surrounding a crash, such as the time of day or if property damage occurred.
+Thus, we ended up not going with the alternative dataset for this paper, but there are variables in the alternative dataset that may motivate future research on this subject.
+(EDIT TO MAKE MORE CLEAR)
+
+The paper uses the R programming language [@citeR] to analyze the dataset.
+The tidyverse package was used to simulate the dataset.
+Also, the tidyverse [@citetidyverse], arrow [CITE] and opendatatoronto [@citeopendatatoronto] packages were used to download the Victims of Crime dataset.
+Then, the tidyverse [@citetidyverse] package was used to clean the raw dataset and generate tests.
+The testthat package [CITE THIS] was used to create tests for our cleaned dataset.
+Rstanarm [CITE], Arrow [CITE], and bayestestR [CITE] were used to create and test the model.
+Finally, ggplot2 [@citeggplot2], tidyverse [@citetidyverse], knitr [@citeknitr] and scales [@citescales] packages were used to create the tables and graphs to display the results.
+[edit this paragraph]
+
+## Data Cleaning (MAYBE INCLUDE IN OTHER PLACE? )
+
+Explain how we cleaned fatalities to a variable rather than whatever
+
+The variables of interest in the paper are CrimeType (named SUBTYPE in the original dataset), which categories the type of crime into the four areas of Assault, Sexual Violation, Robbery and Other; AssaultType (named ASSAULT_SUBTYPE in the original dataset), which specifies assault on peace officers into the subtypes noted in @sec-introduction; Sex (named SEX in the original dataset) - which is broken down into Male, Female and Unknown, where Unknown means the victim’s sex is not known by the Toronto Police Service; and Count (named COUNT in the original dataset), which counts the number of identified victims who share the same demographic characteristics previously.
+Each entry in the dataset does not represent a unique person but notes the number of people who share the same characteristics, such as Sex and CrimeType.
+Other characteristics such as age group, reported year and age cohort were not variables of interest for the study and were removed from the cleaned dataset.
+Using knitr [@citeknitr], the first 10 lines of the cleaned dataset is shown in @sec-appendix-sample under @tbl-head.
+In addition, summaries of the Sex and Year variables are displayed in @sec-appendix-summary_statistics.
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
 ```{r}
@@
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
 ```
@@
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
 ```
@@
 
+### Hour
+
 Add graphs, tables and text.
@@
 Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+
+### Type of Crash
+
+### Automobile
+
+### Motorcycle
+
+### Passenger
+
+### Pedestrian
+
+### Police Division
+Fixed effects
+
+### Year 
+Fixed effects
+
+### Interaction Terms
+Why you have interaction between hour and all these things
 
@@
 
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
 ```
@@
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+#modelsummary::modelsummary(
+ # list(
+  #  "First model" = first_model
+  #),
+  #statistic = "mad",
+  #fmt = 2
+#)
 ```
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+In we implement a posterior predictive check. This shows...
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In we compare the posterior with the prior. This shows...
 
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
+#pp_check(first_model) +
+  #theme_classic() +
+ # theme(legend.position = "bottom")
 
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
+#posterior_vs_prior(first_model) +
+ # theme_minimal() +
+  #scale_color_brewer(palette = "Set1") +
+ # theme(legend.position = "bottom") +
+ # coord_flip()
 ```
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+is a trace plot. It shows... This suggests...
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+is a Rhat plot. It shows... This suggests...
 
@@
 
-plot(first_model, "trace")
+#plot(first_model, "trace")
 
-plot(first_model, "rhat")
+#plot(first_model, "rhat")
 ```

### 2024-12-02 scripts/05-model_data.R
updated factors and deleted base model

@@ -19,36 +19,24 @@ analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 ### Model data ####
 
 # Convert variables to factors
-analysis_data$month <- factor(analysis_data$month)
+analysis_data$hour <- factor(analysis_data$hour)
 analysis_data$year <- factor(analysis_data$year)
 analysis_data$police_division <- factor(analysis_data$police_division)
 
-# Get the reduced dataset of only 1000 randomly selected data entries
-motor_fatality_reduced_data <- slice_sample(analysis_data, n = 2000)
+# Sample a reduced dataset using Stratified Sampling 
+motor_fatality_reduced_data <- analysis_data %>%
+  group_by(fatalities) %>%
+  slice_sample(n = 2000 / n_distinct(analysis_data$fatalities)) %>%
+  ungroup()
 
 set.seed(420)
-
-# Base Model
-motor_fatality_base_model<-
-  stan_glm(
-    fatalities ~ hour + injury_collision + fail_to_remain_collision + property_damage_collision + automobile + 
-      motorcycle + passenger + bicycle + pedestrian,
-    data = motor_fatality_reduced_data,
-    family = binomial(link = "logit"),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = 
-      normal(location = 0, scale = 2.5, autoscale = TRUE),
-    seed = 420,
-    iter = 2000,
-    init = "0"
-  )
 
 # Extension Model with Interaction and Fixed Effects
 
 motor_fatality_prediction_model <-
   stan_glm(
     fatalities ~ hour + injury_collision + fail_to_remain_collision + property_damage_collision + automobile + 
-      motorcycle + passenger + bicycle + pedestrian + police_division + year + hour*automobile + hour*pedestrian + hour* motorcycle + hour*passenger + hour*bicycle,
+      motorcycle + passenger + bicycle + pedestrian + police_division + year,
     data = motor_fatality_reduced_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -61,13 +49,7 @@ motor_fatality_prediction_model <-
 
 
 #### Save models ####
-summary(motor_fatality_base_model)
 summary(motor_fatality_prediction_model)
-
-saveRDS(
-  motor_fatality_base_model,
[…]

### 2024-12-02 scripts/05-model_data.R
readded interaction terms that are more relevant

@@ -36,7 +36,7 @@ set.seed(420)
 motor_fatality_prediction_model <-
   stan_glm(
     fatalities ~ hour + injury_collision + fail_to_remain_collision + property_damage_collision + automobile + 
-      motorcycle + passenger + bicycle + pedestrian + police_division + year,
+      motorcycle + passenger + bicycle + pedestrian + police_division + year + hour*motorcycle + hour*pedestrian,
     data = motor_fatality_reduced_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-12-02 paper/paper.qmd
added draft of methodology section

@@
 analysis_data <- read_parquet(file = here("data/02-analysis_data/analysis_data.parquet"), show_col_types = FALSE)
+raw_data <- read_parquet(file = here("data/01-raw_data/raw_data.parquet"), show_col_types = FALSE)
 
@@
 The use of statistical modeling on increasingly available vehicle collision data presents an opportunity to develop a nuanced understanding of what factors increases the likelihood for a fatality to occur in an accident.
-This paper uses the Toronto Police Service's  Annual Statistical Report from Open Data Toronto to analyze what factors are most responsible in predicting if a fatality occurs in a car crash.
+This paper uses the Toronto Police Service's  Annual Statistical Report from Open Data Toronto to analyze what factors are most responsible in predicting if a fatality occurs in a MVC.
 
-The estimand of interest is the log-adjusted probability of a fatality occuring in a car crash.
+The estimand of interest is the log-adjusted probability of a fatality occuring in a MVC.
 Specifically, we aim to quanitfy how specific environmental factors increase or decrease the likelihood of a fatality.
@@
 The dataset is updated annually, is open data, and can be used if an attribution statement @sec-appendix-attribution and is properly cited [@tphlicense].
+Each entry in the dataset represents a singular vehicular accident and records all MVCs from 2015.
 
@@
 
-## Data Cleaning (MAYBE INCLUDE IN OTHER PLACE? )
-
-Explain how we cleaned fatalities to a variable rather than whatever
-
-The variables of interest in the paper are CrimeType (named SUBTYPE in the original dataset), which categories the type of crime into the four areas of Assault, Sexual Violation, Robbery and Other; AssaultType (named ASSAULT_SUBTYPE in the original dataset), which specifies assault on peace officers into the subtypes noted in @sec-introduction; Sex (named SEX in the original dataset) - which is broken down into Male, Female and Unknown, where Unknown means the victim’s sex is not known by the Toronto Police Service; and Count (named COUNT in the original dataset), which counts the number of identified victims who share the same demographic characteristics previously.
-Each entry in the dataset does not represent a unique person but notes the number of people who share the same characteristics, such as Sex and CrimeType.
-Other characteristics such as age group, reported year and age cohort were not variables of interest for the study and were removed from the cleaned dataset.
-Using knitr [@citeknitr], the first 10 lines of the cleaned dataset is shown in @sec-appendix-sample under @tbl-head.
-In addition, summaries of the Sex and Year variables are displayed in @sec-appendix-summary_statistics.
-
 ## Measurement
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+Transforming a real-life Motor Vehicle Collision to an entry in the dataset is a well-documented process by the Toronto Police Service.
+For insurance purposes, the Toronto Police Service requires drivers to fill out the Motor Vehicle Collision Report for any collisions that occur in Toronto (CITE).
+Drivers required to fill out a motor vehicle collision report if the combined damage is more than $2000, if someone is injured, if a criminal act such as a DUI occurs, or if a pedestrian is involved in the accident (CITE).
+These reports are retained for six years by the Toronto Police Service, with the exception of collisions resulting in a fatality, which are retained indefinitely.
+The form ensures documentation of collision characteristics, location, road condition, and the extent of damages, systematically recording the characteristics of each crash for further criminal investigation and data analysis.
+
+For every collision, basic facts such as the location, time, and date of the collision is recorded through the Motor Vehicle Collision Report.
+Majority if not all the factors recorded in the dataset are all objective measurements regarding the specific details such as if a motorcycle was involved in the collision or if the collision resulted in property damage.
+All of these details are recorded in the Motor Vehicle Collision Report for all vehicular collisions and are entered into the dataset.
+However, while the Motor Vehicle Collision Report logs characteristics such as environmental conditions, alcohol involvement, or fatigue, the data set does not include them due to inconsistent data measurement techniques.
+Moreover, personal details such as the driver's age are not included to protect the driver's privacy.
 
@@
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+The main outcome variable records the number of fatalities for each MVC.
+According to the dataset, a fatal collision occurs when an individual's injuries from a collision results in a fatality within 30 days.
+Fatal collisions excludes occurences on private property, ones related to suddend eath prior to collision, such as suicide, and where the individuals has died more than 30 days after the collision.
+However, because we are more interested in predicting the probability that a fatality occurs than the number of fatalities, we transformed the variable that distinguishes collisions between if the collision resulted in any fatalities and those without fatalities.
+In the raw dataset, if there were no fatalities, the entry was recorded as NA, but if there were fatalities, then the number of fatalities were recorded.
+However, we transformed the dataset that all if a fatality occurred then fatalities indicates `1` and if there were no fatalities then the fatalities column records a '0'.
+The distribution of the raw dataset is shown in @fig-fatalities-raw.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-fatalities-raw
+#| fig-cap: Distribution of Fatalities in the Raw Dataset  
 #| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Plot distribution of Fatalies in the Raw Dataset 
+ggplot(raw_data %>% mutate(FATALITIES = replace_na(FATALITIES, 0)), 
+       aes(x = FATALITIES)) +
+  geom_bar(fill = "orange", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Fatalities",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+```
+However, after the transformation, the outcome variable now takes on binary values and the distribution and summary statistics are shown below in @fig-fatalites-cleaned and @tbl-fatalites:
+
+```{r}
+#| label: fig-fatalities-cleaned
+#| fig-cap: Distribution of Fatalities in the Cleaned Dataset   
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Plot distribution of Fatalities
+ggplot(analysis_data, aes(x = fatalities)) +
+  geom_bar(fill = "green", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Fatalities",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
 
@@
 
-Talk more about it.
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-fatalities
+#| tbl-cap: Number of unique high-quality polling organizations
+#| tbl-align: center
+#| tbl-width: 82%
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+# Calculate summary statistics for the 'fatalities' variable
+summary_stats <- analysis_data %>%
+  count(fatalities, name = "Fatalities")
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
 
 ```
-
-Talk way more about it.
-
 ## Predictor variables
@@
 
-Add graphs, tables and text.
+The hour variable indicates the time at which the accident occurred using the 24 hour clock, such that a value of 0 represents 12 AM and 23 represents 11 PM.
+The distribution and summary statistics of the hour variable can be found in @fig-hour and @tbl-hour, respectively.
+@fig-hour highlights that the majority of accidents happen from 9 AM to 7 PM, which is reasonable as these times include rush hour, where the greatest number of people are driving at the same time for work or school.
+However, @tbl-hour also shows that a greater number of fatalities occur at night, which we hypothesize is due to lack of vision or reckless driving.
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+```{r}
+#| label: fig-hour
+#| fig-cap: Distribution of the Hour Variable  
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Plot distribution of Hour Variable
+ggplot(analysis_data, aes(x = hour)) +
+  geom_bar(fill = "yellow", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Hour",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-hour
+#| tbl-cap: The Summary of the Hour 
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for the 'hour' variable
+summary_stats <- analysis_data %>%
+  group_by(hour) %>%
+  summarize(
+    "No Fatalities" = sum(fatalities == 0, na.rm = TRUE),
+    "Fatality Occurred" = sum(fatalities == 1, na.rm = TRUE),
+    Total = n()
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
 
@@
 
+Beyond identifying fatalities, the Motor Vehicle Collision Reports notes if the crash was one of three types: an injury collision, a fail to remain collision, and a property damage collision.
+A personal injury collision occur when an individual involved in a MVC suffers personal injuries.
+Fail to remain collisions are recorded when an individual involved in a collision fail to stop and provide their information at the scene of a collision.
+Property damage collisions occur when an individual has been damaged in a collision or the value of damages is less than $\2000 for all parties.
+The distribution and summary statistics of these three variables can be found in @fig-crashtype and @tbl-crashtype.
+The results in @tbl-crashtype show that crashes that classify under these three categories usually do not lead to death.
+
+```{r}
+#| label: fig-crashtype
+#| fig-cap: Breakdown of Each Collision Type by Fatalities  
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Reshape data to long format for easier plotting
+analysis_data_long <- analysis_data %>%
+  select(fatalities, injury_collision, fail_to_remain_collision, property_damage_collision) %>%
+  gather(key = "collision_type", value = "collision", 
+         injury_collision, fail_to_remain_collision, property_damage_collision)
+
+# Plot the data
+ggplot(analysis_data_long, aes(x = collision_type, fill = factor(fatalities))) +
+  geom_bar(position = "dodge", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Collision Type", y = "Count") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_fill_manual(values = c("0" = "green", "1" = "red"), 
+                    labels = c("0" = "No Fatalities", "1" = "Fatality Occurred")) +
+  scale_x_discrete(labels = c("injury_collision" = "Injury Collision", 
+                              "fail_to_remain_collision" = "Fail to Remain Collision", 
+                              "property_damage_collision" = "Property Damage Collision"))
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-crashtype
+#| tbl-cap: Breakdown of MVCs Into the Different Categories, Broken Down by Fatalities
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Reshape data to long format for easier plotting
+
+# Calculate summary statistics for the 'Collision Type' variable
+summary_stats <- analysis_data %>%
+  mutate(
+    collision_type = case_when(
+      injury_collision == 1 ~ "Injury Collision",
+      fail_to_remain_collision == 1 ~ "Fail to Remain Collision",
+      property_damage_collision == 1 ~ "Property Damage Collision",
+      TRUE ~ "Not Applicable"
+    )
+  ) %>%
+  group_by(collision_type) %>%
+  summarize(
+    "No Fatalities" = sum(fatalities == 0, na.rm = TRUE),
+    "Fatality Occurred" = sum(fatalities == 1, na.rm = TRUE),
+    Total = n()
+  ) %>%
+  # Add a row for total at the bottom
+  bind_rows(
+    summarize(
+      analysis_data,
+      collision_type = "Total",
+      "No Fatalities" = sum(fatalities == 0, na.rm = TRUE),
+      "Fatality Occurred" = sum(fatalities == 1, na.rm = TRUE),
+      Total = n()
+    )
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+
 ### Automobile
+
+The automobile variable is a indicator variable to show if a collision involved a person in an automobile.
+In the raw dataset, the variable was labeled as Yes, No, None or N/R (Not Recorded).
+We labelled Not Recorded as NA because there is no reliable way of characterizing the variable.
+Further, we labeled No and None as 0 and Yes as 1, where 1 represents that an automobile was involved and 0 represents that an automobile was not, such as a crash between two motorcycles.
+We also employed this method for the following variables: motorcycle, passenger, and pedestrian.
+
+@fig-automobile and @tbl-automobile shows that 588 of 608 deaths happened when an automobile was involved, which is not surprising given majority of vehicles on the road are cars.
+
+```{r}
+#| label: fig-automobile
+#| fig-cap: Distribution of the Automobile Variable  
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Plot distribution of Hour Variable
+ggplot(analysis_data, aes(x = automobile)) +
+  geom_bar(fill = "purple", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Automobile",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-automobile
+#| tbl-cap: Summary of the Automobile Variable 
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for the 'hour' variable
+summary_stats <- analysis_data %>%
+  group_by(automobile) %>%
+  summarize(
+    "No Fatalities" = sum(fatalities == 0, na.rm = TRUE),
+    "Fatality Occurred" = sum(fatalities == 1, na.rm = TRUE),
+    Total = n()
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
 
@@
 
+The motorcycle variable is another indicator variable to show that whether the collision involved a person in a motorcycle.
+1 represents that a motorcycle was involved and 0 represents that a motorcycle was not involved in the crash.
+@fig-motorcycle and @tbl-motorcycle shows the distribution and summary of the motorcycle variable, respectively.
+@tbl-motorcycle shows that only 75 vehicular deaths involved a motorcycle.
+
+```{r}
+#| label: fig-automobile
+#| fig-cap: Distribution of the Automobile Variable  
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Plot distribution of Hour Variable
+ggplot(analysis_data, aes(x = motorcycle)) +
+  geom_bar(fill = "orange", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Motorcycle",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-motorcycle
+#| tbl-cap: Summary of the Automobile Variable 
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for the 'hour' variable
+summary_stats <- analysis_data %>%
+  group_by(motorcycle) %>%
+  summarize(
+    "No Fatalities" = sum(fatalities == 0, na.rm = TRUE),
+    "Fatality Occurred" = sum(fatalities == 1, na.rm = TRUE),
+    Total = n()
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
 ### Passenger
 
+The passenger variable is an indicator variable that highlights if the collision involved a passenger in a motor vehicle.
+1 represents there was a passenger and 0 shows that there was not a passenger involved.
+@fig-passenger and @tbl-passenger shows the distribution and summary of the passenger variable, respectively.
+@tbl-passenger shows that 177 vehicular deaths involved a passenger.
+
+```{r}
+#| label: fig-passenger
+#| fig-cap: Distribution of the Passenger Variable  
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Plot distribution of Hour Variable
+ggplot(analysis_data, aes(x = passenger)) +
+  geom_bar(fill = "blue", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Passenger",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-passenger
+#| tbl-cap: Summary of the Passenger Variable 
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for the 'hour' variable
+summary_stats <- analysis_data %>%
+  group_by(passenger) %>%
+  summarize(
+    "No Fatalities" = sum(fatalities == 0, na.rm = TRUE),
+    "Fatality Occurred" = sum(fatalities == 1, na.rm = TRUE),
+    Total = n()
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
 ### Pedestrian
 
+The pedestrian variable is an indicator variable that highlights if the collision involved a pedestrian.
+1 represents a pedestrian was involved and 0 shows that there was no pedestrian.
+@fig-pedestrian and @tbl-pedestrian shows the distribution and summary of the pedestrian variable, respectively.
+@tbl-pedestrian highlights that of 606 deaths, 342 deaths involved pedestrians, which is a significant percentage.
+
+```{r}
+#| label: fig-pedestrian
+#| fig-cap: Distribution of the Pedestrian Variable  
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Plot distribution of Hour Variable
+ggplot(analysis_data, aes(x = pedestrian)) +
+  geom_bar(fill = "red", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Pedestrian",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-pedestrian
+#| tbl-cap: Summary of the Pedestrian Variable 
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for the 'hour' variable
+summary_stats <- analysis_data %>%
+  group_by(pedestrian) %>%
+  summarize(
+    "No Fatalities" = sum(fatalities == 0, na.rm = TRUE),
+    "Fatality Occurred" = sum(fatalities == 1, na.rm = TRUE),
+    Total = n()
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
 ### Police Division
-Fixed effects
+
+The Police Division variable represents the police division where the collision occured.
+The paper plans to include the police division variable as a general proxy for location and account if certain areas are more susceptible to crashes than others.
+@fig-division and @tbl-division shows the distirbution and breakdown of MVCs among police departments.Based on the @tbl-division, there seems to be no discernible pattern but D41 and D42 have the highest number of vehicular fatalities at 63 and 66, respectively.
+
+```{r}
+#| label: fig-division
+#| fig-cap: Distribution of the Police Division
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Plot distribution of Hour Variable
+ggplot(analysis_data, aes(x = police_division)) +
+  geom_bar(fill = "magenta", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Police Division",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-division
+#| tbl-cap: Summary of the Police Divison Variable 
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for the 'hour' variable
+summary_stats <- analysis_data %>%
+  group_by(police_division) %>%
+  summarize(
+    "No Fatalities" = sum(fatalities == 0, na.rm = TRUE),
+    "Fatality Occurred" = sum(fatalities == 1, na.rm = TRUE),
+    Total = n()
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
 
 ### Year
-Fixed effects
+
+The year variable shows the number of MVCs per year.
+@fig-year and @tbl-year shows the number of MVCs over time.
+Looking at @fig-year and @tbl-year there is a noticeable dip in MVCs in 2020 and 2021 due to COVID-19 but MVC levels have not hit their 2019 peaks most likely due to people in Toronto not driving as much as before.
+However, idea of fewer drivers is a hypothesis and needs further research.
+
+```{r}
+#| label: fig-year
+#| fig-cap: Number of MVCs per Year
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+year_count <- analysis_data %>%
+  count(year, name = "count")
+
+# Plot distribution of Year Variable
+ggplot(year_count, aes(x = year, y = count, group = 1)) +
+  geom_line(color = "black", alpha = 0.7) +
+  theme_minimal() +
+  labs(x = "Year",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-year
+#| tbl-cap: Summary of Number of MVCs by Year
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for the 'hour' variable
+summary_stats <- analysis_data %>%
+  group_by(year) %>%
+  summarize(
+    "No Fatalities" = sum(fatalities == 0, na.rm = TRUE),
+    "Fatality Occurred" = sum(fatalities == 1, na.rm = TRUE),
+    Total = n()
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
 
@@
 # Model
+
+The study aims to predict the likelihood of fatality in a car crash. The model was implemeneted using the rstatarm package in R, leveraging its framework for Baysian regression modelling.
 
@@
 
-#modelsummary::modelsummary(
+modelsummary::modelsummary(
  # list(

### 2024-12-02 paper/paper.qmd
added residual parameters and progressed paper

@@
 ```{r}
-#| label: fig-automobile
+#| label: fig-motorcycle
 #| fig-cap: Distribution of the Automobile Variable  
@@
 
-### Interaction Terms
+### Relevant Interaction Terms
 Why you have interaction between hour and all these things
@@
 
-The study aims to predict the likelihood of fatality in a car crash. The model was implemeneted using the rstatarm package in R, leveraging its framework for Baysian regression modelling.
+## Model Set-Up and Justification
 
-The goal of our modelling strategy is twofold. Firstly,...
+For our analysis, we employ a Bayesian Logistic Linear Model to forecast the likelihood of a fatality in a car crash.
+This approach allows us to capture known variations between years, police divisions, and the characteristics of a crash, such as the involvement of a motorcycle or if someone was injured.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The first step in the process involved selecting a reliable dataset for model development.
+We utilized high-quality vehicle collision data gathered by the Toronto Police Service.
+We first excluded all cases in our predictor variables that had null values.
+In addition, we transformed fatalities into a binary response variable.
+Let $Y_i$ represent the fatality outcome for observation $i$, where $Y_i =1$ indicates a fatality occurred and $Y_i=0$ indicates no fatality.
+The probability of fatality is modeled through a Bernoulli distribution:
 
-## Model set-up
+$$Y_i = 
+\begin{cases} 
+0 & \text{if no fatality occurred (original classification: N/A)} \\ 
+1 & \text{if any fatality occurred (original classifications: 1, 2, or 3 fatalities)}
+\end{cases}$$
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+\begin{align}S_i \sim \text{Bernoulli}(p_i)\end{align}
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+\begin{equation}
+\text{logit}(p_i) = \beta_{0} 
++ \sum_{j=1}^{n} \beta_{\text{hour}_j} x_{\text{hour}_{ij}} 
++ \beta_{2} x_{\text{InjuryCollision}_i} 
++ \beta_{3} x_{\text{FTRCollision}_i} 
++ \beta_{4} x_{\text{PDCollision}_i} 
++ \beta_{5} x_{\text{automobile}_i} 
++ \beta_{6} x_{\text{motorcycle}_i} 
++ \beta_{7} x_{\text{passenger}_i} 
++ \beta_{8} x_{\text{pedestrian}_i} 
++ \sum_{z=1}^{a} \beta_{\text{PoliceDivision}_z} x_{\text{PoliceDivision}_{iz}} 
++ \sum_{m=1}^{a} \beta_{\text{year}_m} x_{\text{year}_{im}} 
++ \sum_{j=1}^{n} \beta_{\text{hour}_j \times \text{motorcycle}} x_{\text{hour}_{ij}} \cdot x_{\text{motorcycle}_i} 
++ \sum_{j=1}^{n} \beta_{\text{hour}_j \times \text{pedestrian}} x_{\text{hour}_{ij}} \cdot x_{\text{pedestrian}_i} 
++ \epsilon_{i}
+\label{eq:logit}
+\end{equation}
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+In equation \ref{eq:logit}, each $\beta$ represents a coefficient determined through regression analysis.
+The variables chosen for the model are hour, the different types of collisions, automobile, motorcycle, passenger, pedestrian, year, and police divsion.
+Each predictor variable was selected based on its significance in vehicle fatality prediction analysis.
+The different types of collisions
 
-### Model justification
+Hour5 and hour9 motorcycle dropped for zero effects
+fix for location effects
+temporal effects
+interaction terms fix for how different parts of the hour best affect (bicycle not included as had a precise null effect and prevent overfitting)
+
+ The identity of the polling organization is important to our model, as each pollster may exhibit unique biases.
+Including Pollster as a fixed effect allows us to account for these variations without introducing unnecessary complexity.
+Population Type distinguishes among different groups surveyed (e.g., voters, likely voters) and helps in capturing generalizability.
+Categorizing polls as either recent or older ensures that more recent polls are better predictors of voting behavior closer to the election.
+$Y_i$ denotes the predicted popular vote percentage for Kamala Harris in the $i$-th poll.
+$\epsilon_i$ is the Gaussian-distributed error term, accounting for residual variation in the model.
+
+To enhance the model, Bayesian priors were applied, introducing regularization and incorporating plausible ranges grounded.
+For the coefficient priors $\beta$, a normal distribution with a mean of 0 and a scale of 2.5 (autoscaled) was chosen to provide flexibility while mitigating overfitting.
+Similarly, the intercept uses a normal prior with a mean of 0 and a scale of 2.5 to stabilize model estimates.
+For the error term (sigma), an exponential prior with a rate of 1 was selected to constrain the residuals, aligning with Gaussian assumptions.
+
+The model was implemented in R [@citeR] using the `rstanarm` package, which offers an accessible interface for Bayesian generalized linear models (GLMs), allowing specification of priors and customization of model parameters.
+Once the logistic regression model for predicting the popular vote is developed, we apply the `predict()` function in R [@citeR] to generate popular vote percentage predictions for Harris.
+This prediction is then added to the dataset as a new variable (`predicted_pct`) for further analysis.
+The results are saved in a CSV file (`vote_for_Harris_predictions.csv`), enabling an examination of potential outcomes in the vote.
+This approach facilitates an understanding of vote distributions at a national level and provides a foundation for forecasting electoral outcomes based on polling data.
+
+## Model Assumptions and Limitations
 
@@
 
-modelsummary::modelsummary(
+#modelsummary::modelsummary(
  # list(
@@
 
-## Third discussion point
+## Implications
 
-## Weaknesses and next steps
+## Weaknesses and Next Steps
 

### 2024-12-02 scripts/05-model_data.R

@@ -42,6 +42,7 @@ motor_fatality_prediction_model <-
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = 
       normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_aux = exponential(rate = 1, autoscale = TRUE),
     seed = 420,
     iter = 2000,
     init = "0"

### 2024-12-02 paper/paper.qmd
added results section

@@
 library(kableExtra)
+library(sjPlot)
 
@@
 \text{logit}(p_i) = \beta_{0} 
-+ \sum_{j=1}^{n} \beta_{\text{hour}_j} x_{\text{hour}_{ij}} 
-+ \beta_{2} x_{\text{InjuryCollision}_i} 
-+ \beta_{3} x_{\text{FTRCollision}_i} 
-+ \beta_{4} x_{\text{PDCollision}_i} 
-+ \beta_{5} x_{\text{automobile}_i} 
-+ \beta_{6} x_{\text{motorcycle}_i} 
-+ \beta_{7} x_{\text{passenger}_i} 
-+ \beta_{8} x_{\text{pedestrian}_i} 
-+ \sum_{z=1}^{a} \beta_{\text{PoliceDivision}_z} x_{\text{PoliceDivision}_{iz}} 
-+ \sum_{m=1}^{a} \beta_{\text{year}_m} x_{\text{year}_{im}} 
-+ \sum_{j=1}^{n} \beta_{\text{hour}_j \times \text{motorcycle}} x_{\text{hour}_{ij}} \cdot x_{\text{motorcycle}_i} 
-+ \sum_{j=1}^{n} \beta_{\text{hour}_j \times \text{pedestrian}} x_{\text{hour}_{ij}} \cdot x_{\text{pedestrian}_i} 
++ \beta_{1} x_{\text{hour}_{i}} 
++ \beta_{2} x_{\text{InjuryCollision}_{i}} 
++ \beta_{3} x_{\text{FTRCollision}_{i}} 
++ \beta_{4} x_{\text{PDCollision}_{i}} 
++ \beta_{5} x_{\text{automobile}_{i}} 
++ \beta_{6} x_{\text{motorcycle}_{i}} 
++ \beta_{7} x_{\text{passenger}_{i}} 
++ \beta_{8} x_{\text{pedestrian}_{i}} 
++ \beta_{9} x_{\text{PoliceDivision}_{i}} 
++ \beta_{10} x_{\text{year}_{im}} 
++ \beta_{11} x_{\text{hour}_{i}} \cdot x_{\text{motorcycle}_{i}} 
++ \beta_{12} x_{\text{hour}_{i}} \cdot x_{\text{pedestrian}_{i}} 
 + \epsilon_{i}
@@
 In equation \ref{eq:logit}, each $\beta$ represents a coefficient determined through regression analysis.
-The variables chosen for the model are hour, the different types of collisions, automobile, motorcycle, passenger, pedestrian, year, and police divsion.
+The variables chosen for the model are hour, the different types of collisions, automobile, motorcycle, passenger, pedestrian, year, and police division.
 Each predictor variable was selected based on its significance in vehicle fatality prediction analysis.
-The different types of collisions
-
-Hour5 and hour9 motorcycle dropped for zero effects
-fix for location effects
-temporal effects
-interaction terms fix for how different parts of the hour best affect (bicycle not included as had a precise null effect and prevent overfitting)
-
- The identity of the polling organization is important to our model, as each pollster may exhibit unique biases.
-Including Pollster as a fixed effect allows us to account for these variations without introducing unnecessary complexity.
-Population Type distinguishes among different groups surveyed (e.g., voters, likely voters) and helps in capturing generalizability.
-Categorizing polls as either recent or older ensures that more recent polls are better predictors of voting behavior closer to the election.
-$Y_i$ denotes the predicted popular vote percentage for Kamala Harris in the $i$-th poll.
-$\epsilon_i$ is the Gaussian-distributed error term, accounting for residual variation in the model.
+Temporal variables such as hour and year are introduced as fixed effects to account for biases across time without introducing unnecessary complexity.
+Similarly, location introduces biases as some neighborhoods may be more susceptible to vehicle accidents than others, prompting the inclusion of police department as a fixed effect and a proxy for location.
+Furthermore, depending on the time of day, motorcyclists and pedestrians are more susceptible to fatal vehicle accidents than during the day.
+Thus, $\beta_{11}$ and $\beta_{12}$ represents the coefficient of an interaction term between motorcycle and pedestrian with the categorical variable hour, respectively, to account for potential variations in risk patterns associated with different times of the day.
+These interaction terms allow the model to capture the increased vulnerability of motorcyclists and pedestrians during specific hours, such as nighttime or early morning, when visibility and traffic conditions may differ significantly, thereby improving the precision and interpretability of the predictions.
+In addition, any variables, especially the various interaction terms, will be omitted if they display high multicolinearity or insignificance.
+Finally, $\epsilon_i$ is the Gaussian-distributed error term, accounting for residual variation in the model.
 
@@
 The model was implemented in R [@citeR] using the `rstanarm` package, which offers an accessible interface for Bayesian generalized linear models (GLMs), allowing specification of priors and customization of model parameters.
-Once the logistic regression model for predicting the popular vote is developed, we apply the `predict()` function in R [@citeR] to generate popular vote percentage predictions for Harris.
-This prediction is then added to the dataset as a new variable (`predicted_pct`) for further analysis.
-The results are saved in a CSV file (`vote_for_Harris_predictions.csv`), enabling an examination of potential outcomes in the vote.
-This approach facilitates an understanding of vote distributions at a national level and provides a foundation for forecasting electoral outcomes based on polling data.
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+The model assumes independence of observations, meaning that each observation's outcome is not influenced by others.
+However, this assumption may be partially violated in the context of vehicle accidents, as incidents occurring within close temporal or spatial proximity may share common influencing factors, such as weather, road conditions, or traffic patterns.
+While the inclusion of temporal and spatial variables like hour and police division aims to mitigate such dependencies, it may not fully account for potential clustering effects.
+As a result, the model could underestimate or overestimate the significance of some predictors, impacting the robustness of its conclusions.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+Another important limitation is the potential for omitted variable bias and unmeasured confounding factors.
+For example, the model does not include variables like road infrastructure quality, driver impairment (e.g., alcohol or drug use), or real-time weather conditions, all of which could influence accident outcomes.
+Additionally, selection bias is a concern, as the dataset may disproportionately represent severe accidents reported for insurance claims while omitting less severe incidents.
+This could skew the model’s predictions toward more severe outcomes and limit its generalizability.
+Addressing these limitations would require incorporating additional data sources and applying techniques to account for potential dependencies and missing information.
+Additionally, the model relies on reported data, which may include potential selection bias, as minor accidents are less likely to be documented.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+To assess model reliability, we examined several key metrics.
+Convergence metrics, such as Rhat values, were very close to 1 for all parameters, indicating strong convergence.
+Additionally, the effective sample size $\text{n_{eff}}$ was high across all parameters, suggesting low autocorrelation and contributing to model stability.
+See more details of our model diagnostics here: (INSERT REFERENCE)
 
@@
 #| echo: false
-#| eval: true
+#| message: false
 #| warning: false
-#| message: false
+#| tbl-align: center
+#| label: tbl-modelresults
+#| tbl-cap: "Summary for Motor Vehicle Collision Fatality Prediction Model"
 
-library(rstanarm)
-
+modelsummary(
+  prediction_model,
+  output = "kableExtra",
+) 
 ```
+@tbl-modelresults presents the estimated coefficients for the predictors in our GLM model.
+These coefficients fit into the GLM equation, \ref{eq:logit}, allowing us to interpret the impact of each predictor on the likelihood of fatality in a vehicular accident.
+Most notably, motorcyclist (18.723) and pedestrian (7.403) collisions exhibit substantial positive associations with fatality occurance, which is in line with expectations.
+Moreover, motorcyclist accidents during rush hour (hour 16 and 17) have the highest coefficients at 32.688 and 21.926, respectively, and during the late evening, such as at 11:00 PM (34.365), indicating that a collision involving a motorcycle during rush hour or late evening is around 77\% likely to result in a fatality.
+We also find similar results with pedestrians as well.
+The type of collision variables all have a significant negative effect on the likelihood of death, as well.However, police division and year don't seem to have a significant impact on the likelihood of death.
 
@@
 #| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| message: false
 #| warning: false
+#| fig-align: center
+#| label: fig-coefficient-estimates
+#| fig-cap: "Coefficient Estimates for Predictors"
 
-#modelsummary::modelsummary(
- # list(
-  #  "First model" = first_model
-  #),
-  #statistic = "mad",
-  #fmt = 2
-#)
+# Tidy Model Summary Data
+coefficients <- broom.mixed::tidy(prediction_model, conf.int = TRUE) %>%
+  mutate(across(c(estimate, std.error, conf.low, conf.high), round, 2))
+
+# Make Plot
+coefficients %>%
+  ggplot(aes(estimate, term)) + 
+  geom_point() + 
+  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) + 
+  labs(x = "Estimate", y = "Predictor") +
+  theme(axis.text.y = element_text(size = 7))
 ```
+
+@fig-coefficient-estimates represents the model coefficients, with error bars indicating the confidence interval for each estimate.
+While most variables are relatively in line, it seems that the automobile and motorcycle variables demonstrate high variability in its estimates.
 

### 2024-12-02 paper/paper.qmd
added some alignment elements for the equation

@@
 
-\begin{equation}
-\text{logit}(p_i) = \beta_{0} 
+\newpage
+
+\begin{align}
+\text{logit}(p_i) &= \beta_{0} 
 + \beta_{1} x_{\text{hour}_{i}} 
@@
 + \beta_{3} x_{\text{FTRCollision}_{i}} 
-+ \beta_{4} x_{\text{PDCollision}_{i}} 
-+ \beta_{5} x_{\text{automobile}_{i}} 
++ \beta_{4} x_{\text{PDCollision}_{i}} \nonumber \\ 
+&\quad + \beta_{5} x_{\text{automobile}_{i}} 
 + \beta_{6} x_{\text{motorcycle}_{i}} 
 + \beta_{7} x_{\text{passenger}_{i}} 
-+ \beta_{8} x_{\text{pedestrian}_{i}} 
-+ \beta_{9} x_{\text{PoliceDivision}_{i}} 
-+ \beta_{10} x_{\text{year}_{im}} 
-+ \beta_{11} x_{\text{hour}_{i}} \cdot x_{\text{motorcycle}_{i}} 
++ \beta_{8} x_{\text{pedestrian}_{i}} \nonumber \\ 
+&\quad + \beta_{9} x_{\text{PoliceDivision}_{i}} 
++ \beta_{10} x_{\text{year}_{im}} \nonumber \\ 
+&\quad + \beta_{11} x_{\text{hour}_{i}} \cdot x_{\text{motorcycle}_{i}} 
 + \beta_{12} x_{\text{hour}_{i}} \cdot x_{\text{pedestrian}_{i}} 
@@
 \label{eq:logit}
-\end{equation}
+\end{align}
 
@@
 Convergence metrics, such as Rhat values, were very close to 1 for all parameters, indicating strong convergence.
-Additionally, the effective sample size $\text{n_{eff}}$ was high across all parameters, suggesting low autocorrelation and contributing to model stability.
+Additionally, the effective sample size $n_{\text{eff}}$ was high across all parameters, suggesting low autocorrelation and contributing to model stability.
 See more details of our model diagnostics here: (INSERT REFERENCE)

### 2024-12-03 scripts/05-model_data.R
added sketches

@@ -24,10 +24,10 @@ analysis_data$year <- factor(analysis_data$year)
 analysis_data$police_division <- factor(analysis_data$police_division)
 
 # Sample a reduced dataset using Stratified Sampling 
-motor_fatality_reduced_data <- analysis_data %>%
-  group_by(fatalities) %>%
-  slice_sample(n = 2000 / n_distinct(analysis_data$fatalities)) %>%
-  ungroup()
+#motor_fatality_reduced_data <- analysis_data %>%
+  #group_by(fatalities) %>%
+  #slice_sample(n = 2000 / n_distinct(analysis_data$fatalities)) %>%
+  #ungroup()
 
 set.seed(420)
 
@@ -37,7 +37,7 @@ motor_fatality_prediction_model <-
   stan_glm(
     fatalities ~ hour + injury_collision + fail_to_remain_collision + property_damage_collision + automobile + 
       motorcycle + passenger + bicycle + pedestrian + police_division + year + hour*motorcycle + hour*pedestrian,
-    data = motor_fatality_reduced_data,
+    data = analysis_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = 

### 2024-12-03 scripts/05-model_data.R
added errors to model

@@ -24,10 +24,10 @@ analysis_data$year <- factor(analysis_data$year)
 analysis_data$police_division <- factor(analysis_data$police_division)
 
 # Sample a reduced dataset using Stratified Sampling 
-#motor_fatality_reduced_data <- analysis_data %>%
-  #group_by(fatalities) %>%
-  #slice_sample(n = 2000 / n_distinct(analysis_data$fatalities)) %>%
-  #ungroup()
+motor_fatality_reduced_data <- analysis_data %>%
+  group_by(fatalities) %>%
+  slice_sample(n = 2000 / n_distinct(analysis_data$fatalities)) %>%
+  ungroup()
 
 set.seed(420)
 
@@ -37,7 +37,7 @@ motor_fatality_prediction_model <-
   stan_glm(
     fatalities ~ hour + injury_collision + fail_to_remain_collision + property_damage_collision + automobile + 
       motorcycle + passenger + bicycle + pedestrian + police_division + year + hour*motorcycle + hour*pedestrian,
-    data = analysis_data,
+    data = motor_fatality_reduced_data,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = 

### 2024-12-03 scripts/02-download_data.R
updated download script

@@ -30,6 +30,6 @@ library(arrow)
 data <- 
   read_csv("https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/ec53f7b2-769b-4914-91fe-a37ee27a90b3/resource/cb890861-ed20-4862-bb75-b1f9ec1e58dd/download/Traffic%20Collisions%20-%204326.csv")
 #### Save data ####
-write_parquet(data, "data/01-raw_data/raw_data.parquet") 
+write_csv(data, "data/01-raw_data/raw_data.csv") 
 
          

### 2024-12-03 scripts/02-download_data.R
parquet download

@@ -17,19 +17,10 @@ library(arrow)
 
 #### Download data ####
 # Code adapted from OpenDataToronto 
-# Get package
-# package <- show_package("ec53f7b2-769b-4914-91fe-a37ee27a90b3")
 
-# Get all resources for this package
-# resources <- list_package_resources("ec53f7b2-769b-4914-91fe-a37ee27a90b3")
-
-# Identify datastore resources
-# datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
-# Load the first datastore resource as a sample
 data <- 
   read_csv("https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/ec53f7b2-769b-4914-91fe-a37ee27a90b3/resource/cb890861-ed20-4862-bb75-b1f9ec1e58dd/download/Traffic%20Collisions%20-%204326.csv")
 #### Save data ####
-write_csv(data, "data/01-raw_data/raw_data.csv") 
+write_parquet(data, "data/01-raw_data/raw_data.parquet") 
 
          

### 2024-12-03 paper/paper.qmd
added enhancement: model card

@@
 library(sjPlot)
+library(broom.mixed)
+library(gridExtra)
 
@@
 
-what was found (FINISH PARAGRAPH)
+Our analysis identifies key predictors of vehicular accident fatalities, emphasizing heightened risks for motorcyclists (coefficient: 17.036) and pedestrians (coefficient: 6.809), especially during rush hours (hours 16 and 17) and late evenings (11:00 PM), where motorcycle collisions show a fatality likelihood of around 66%.
+Collision type variables, such as injury and property damage collisions, significantly lower fatality risks, while police division and year have minimal impact.
+However, variability in estimates, particularly for motorcyclist and automobile predictors, highlights uncertainty in their precise effects, underscoring the need for targeted interventions during high-risk times and for vulnerable populations.
 
@@
 While the alternative dataset has more explanatory variables simply because more data is collected when someone dies or is seriously injured, this paper's aims to generalize if a fatality is more likely to occur based on the general circumstances surrounding a crash, such as the time of day or if property damage occurred.
-Thus, we ended up not going with the alternative dataset for this paper, but there are variables in the alternative dataset that may motivate future research on this subject.
+In addition, each entry in the alternative dataset represents a killed or injured person, meaning that there exists the possibility for duplicate entries.
+However, each entry in the selected dataset is a unique Motor Vehicle Collision, making it more suitable for the scope of this paper Thus, we ended up not going with the alternative dataset for this paper, but there are variables in the alternative dataset that may motivate future research on this subject.
 (EDIT TO MAKE MORE CLEAR)
@@
 # Calculate summary statistics for the 'fatalities' variable
+
 summary_stats <- analysis_data %>%
-  count(fatalities, name = "Fatalities")
+  summarize(
+    mean = round(mean(fatalities, na.rm = TRUE), 2),
+    median = round(median(fatalities, na.rm = TRUE), 2),
+    min = round(min(fatalities, na.rm = TRUE), 2),
+    max = round(max(fatalities, na.rm = TRUE), 2), 
+    sd = round(sd(fatalities, na.rm = TRUE), 2), 
+    n = n()
+  )
 
@@
 
-### Relevant Interaction Terms
-Why you have interaction between hour and all these things
-
+### Interaction Terms
+
+Due to the compounded effects of the time of hour on motorcyclists and pedestrians in a motor vehicle collision, we controlled for this effect using an interaction term:
+
+#### Hour and Motorcycle
+
+@fig-hour-motorcycle and @tbl-hour-motorcycle shows the visual representation and summary of the interaction term, respectively.
+@fig-hour-motorcycle shows that a significant number of deaths come in the evening from 2 PM to 11 PM.
+
+```{r}
+#| label: fig-hour-motorcycle
+#| fig-cap: Fatality Likelihood by Hour for Motorcyclists
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Create interaction plot data for motorcycle
+interaction_motorcycle <- analysis_data %>%
+  filter(motorcycle == 1) %>%
+  group_by(hour) %>%
+  summarize(fatality_rate_motorcycle = mean(fatalities == 1, na.rm = TRUE))
+
+# Plot the interaction between hour and motorcycle
+ggplot(interaction_motorcycle, aes(x = hour, y = fatality_rate_motorcycle)) +
+  geom_line(color = "red", size = 1.2) +
+  labs(x = "Hour of Day", y = "Motorcycle Fatality Rate") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-hour-motorcycle
+#| tbl-cap: Summary of Fatalities by Hour for Motorcyclists 
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for fatalities by hour for motorcyclists
+summary_stats_motorcycle <- analysis_data %>%
+  filter(motorcycle == 1) %>%
+  group_by(hour) %>%
+  summarize(
+    "Motorcyclist Fatalities" = sum(fatalities == 1, na.rm = TRUE),
+    "Total Motorcyclists" = sum(motorcycle == 1, na.rm = TRUE),
+  )
+
+# Display the summary statistics for motorcyclists in a formatted table
+summary_stats_motorcycle %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
+
+#### Hour and Pedestrian
+
+@fig-hour-pedestrian and @tbl-hour-pedestrian shows the distribution and summary of the interaction variable, respectively.
+@fig-hour-pedestrian, though showing high variability, shows that a lot more pedestrian deaths also happen in the evening.
+
+```{r}
+#| label: fig-hour-pedestrian
+#| fig-cap: Fatality Likelihood by Hour for Pedestrians
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Create interaction plot data for pedestrian
+pedestrian_fatalities <- analysis_data %>%
+  filter(pedestrian == 1) %>%
+  group_by(hour) %>%
+  summarize(pedestrian_fatalities = sum(fatalities == 1, na.rm = TRUE))
+
+# Plot interaction between hour and pedestrian fatalities
+ggplot(pedestrian_fatalities, aes(x = hour, y = pedestrian_fatalities)) +
+  geom_point(color = "red", size = 2) +
+  theme_minimal() +
+  labs(x = "Hour of Day",
+       y = "Pedestrian Fatalities") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-hour-pedestrian
+#| tbl-cap: Summary of Fatalities by Hour for Motorcyclists and Pedestrians
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for fatalities by hour for pedestrians
+summary_stats_pedestrian <- analysis_data %>%
+  filter(pedestrian == 1) %>%
+  group_by(hour) %>%
+  summarize(
+    "Pedestrian Fatalities" = sum(fatalities == 1, na.rm = TRUE),
+    "Total Pedestrians" = sum(pedestrian == 1, na.rm = TRUE),
+  )
+
+# Display the summary statistics for pedestrians in a formatted table
+summary_stats_pedestrian %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
 # Model
@@
 In addition, we transformed fatalities into a binary response variable.
+Moreover, due to the size of the dataset, we conducted a stratified random sampling with equal allocation across the levels of the fatalities variable to ensure each level of the fatalities variable are adequately represented in the model to help the model learn meaningful patterns for that class.
 Let $Y_i$ represent the fatality outcome for observation $i$, where $Y_i =1$ indicates a fatality occurred and $Y_i=0$ indicates no fatality.
@@
 #| tbl-align: center
+#| fig-height: 10
 #| label: tbl-modelresults
@@
 These coefficients fit into the GLM equation, \ref{eq:logit}, allowing us to interpret the impact of each predictor on the likelihood of fatality in a vehicular accident.
-Most notably, motorcyclist (18.723) and pedestrian (7.403) collisions exhibit substantial positive associations with fatality occurance, which is in line with expectations.
-Moreover, motorcyclist accidents during rush hour (hour 16 and 17) have the highest coefficients at 32.688 and 21.926, respectively, and during the late evening, such as at 11:00 PM (34.365), indicating that a collision involving a motorcycle during rush hour or late evening is around 77\% likely to result in a fatality.
+Most notably, motorcyclist (17.036) and pedestrian (6.809) collisions exhibit substantial positive associations with fatality occurrence, which is in line with expectations.
+Moreover, motorcyclist accidents during rush hour (hour 16 and 17) have the highest coefficients at 40.4 and 26.8, respectively, and during the late evening, such as at 11:00 PM (39.365), indicating that a collision involving a motorcycle during rush hour or late evening is around 66\% likely to result in a fatality.
 We also find similar results with pedestrians as well.
@@
 #| label: fig-coefficient-estimates
+#| fig-height: 10
 #| fig-cap: "Coefficient Estimates for Predictors"
@@
 @fig-coefficient-estimates represents the model coefficients, with error bars indicating the confidence interval for each estimate.
-While most variables are relatively in line, it seems that the automobile and motorcycle variables demonstrate high variability in its estimates.
+While most variables are relatively in line, it seems that the interaction terms all have wide confidence intervals, suggesting high variability in its estimates.
 
@@
 
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Implications
-
-## Weaknesses and Next Steps
-
-Weaknesses and next steps should also be included.
+## Contextualizing Role of Motorcyclists and Pedestrians {#sec-discussion-contextualization}
+
+The results of this analysis both validate and challenge several common assumptions about motor vehicle collision (MVC) fatalities.
+Specifically, the high risk faced by motorcyclists, pedestrians, and individuals involved in accidents during rush hours or late evenings aligns with long-held perceptions about these vulnerable groups.
+However, the study also introduces a nuanced understanding of these factors, suggesting deeper complexities that could inform public policy.
+
+### Motorcyclists: The Most Vulnerable Group
+
+In @tbl-modelresults, the coefficient of 17.036 for motorcyclists indicates that they are the most at-risk group in terms of MVC fatalities.
+This result strongly aligns with the widely recognized vulnerability of motorcyclists in traffic accidents.
+Motorcycles, by their very design, offer little protection compared to other vehicles.
+In the event of a crash, motorcyclists are more likely to suffer severe injuries due to the lack of structural safeguards such as airbags, seat belts, or the protective frame that cars provide.
+Additionally, their smaller size makes them less visible to other drivers, increasing the likelihood of collisions.
+This heightened risk is compounded by the fact that motorcycles are often involved in high-speed crashes, particularly on highways or in urban environments with high traffic volumes.
+The coefficient underscores the need for targeted interventions, such as improved rider safety training, better infrastructure (e.g., dedicated motorcycle lanes), and stricter regulations on helmet use.
+
+### Pedestrians
+
+Similarly, shown in @tbl-modelresults, the coefficient of 6.809 for pedestrians reinforces the widely acknowledged dangers pedestrians face in urban settings.
+Pedestrian safety has long been a critical area of concern for urban planners and policymakers, and this result highlights the need for continued focus on this issue.
+Pedestrians are inherently vulnerable in any accident, and the lack of physical barriers between them and vehicles significantly increases their chances of injury or death in a collision.
+Moreover, in many urban areas, crosswalks, sidewalks, and pedestrian signals are either inadequate or poorly maintained, which contributes to pedestrian risk.
+This finding calls for heightened investment in urban infrastructure designed to protect pedestrians, including the expansion of safe crossing points, better traffic signaling, and measures to reduce vehicle speeds in pedestrian-heavy areas.
+It also suggests that policies addressing urban sprawl and public transportation options may help reduce pedestrian fatalities by limiting the need to walk on high-traffic streets.
+
+### Environmental Factors
+
+The analysis also highlights the significant role that environmental factors—specifically, the time of day—play in MVC fatalities.
+During rush hours (particularly between hours 16 and 17) and late evenings (around 11:00 PM), the likelihood of a fatality, especially in motorcycle crashes, increases substantially.
+The 66% fatality likelihood for motorcycle crashes during these times suggests that several environmental and behavioral factors may be at play.
+During rush hours, the density of traffic can lead to higher stress levels among drivers, potentially contributing to reckless driving, lane changes, or a lack of attention to motorcyclists.
+Additionally, the rush-hour period coincides with fatigue, both for drivers who may have been on the road for extended periods and for motorcyclists who may be navigating more complex, congested traffic conditions.
+
+Late evening crashes, on the other hand, are often linked to impaired driving and reduced visibility.
+Fatigue and alcohol impairment increase as the night progresses, leading to a higher likelihood of accidents that involve poor decision-making or delayed reactions.
+In the case of motorcycles, the combination of low visibility, decreased driver attention, and the increased risks of driving under the influence could explain why fatalities are particularly high during late-night hours.
+The study's findings align with common assumptions about these time-based risks, but they also reinforce the importance of implementing targeted interventions, such as stricter enforcement of impaired driving laws and improved lighting and signage to enhance visibility for motorcyclists during late-night hours.
+
+## Role of Protective Factors {#sec-discussion-protective-factors}
+
+The variables "injury_collision" and "property damage collisions" play a crucial role in lowering the likelihood of fatalities in motor vehicle collisions (MVCs).
+These outcomes suggest that, in contrast to fatal accidents, certain collisions are characterized by lower severity, different dynamics, and an overall reduced risk to life.
+By exploring why these variables lower fatality risks, we can better understand the factors that contribute to non-fatal collisions and how interventions might be designed to replicate these protective conditions across all collisions.
+
+Injury collisions often result in less severe damage to both the vehicle and the individuals involved compared to fatal accidents.
+The presence of injuries in these cases indicates that, while there may be harm to the people involved, the force of impact and the nature of the collision may not have been severe enough to result in death.
+Injury collisions may involve safety features, such as airbags, seat belts, or vehicle crumple zones, that significantly mitigate the impact on the occupants, reducing the risk of fatal outcomes.
+Furthermore, when injuries occur, they often prompt quicker medical responses, either by emergency services or through hospitals' ability to treat injuries promptly.
+This quick medical intervention can prevent fatalities by addressing life-threatening injuries before they become fatal.
+Analyzing injury collisions can therefore offer valuable insights into what makes these accidents less deadly and how certain safety features, such as improved emergency response times, advanced vehicle safety systems, and better road infrastructure, can prevent fatalities.
+
+Property damage collisions are another protective factor that lowers the likelihood of fatalities.
+These types of collisions, by definition, involve no physical harm to the individuals involved, but they still represent significant risks of damage to the vehicles and the surrounding environment.
+The lower fatality risk in property damage collisions may be attributed to the collision dynamics being less intense or violent compared to crashes that result in fatalities.
+For example, these types of crashes may occur at lower speeds, with more gradual deceleration, which is less likely to cause life-threatening injuries.
+Additionally, property damage collisions may be associated with specific accident scenarios where vehicles are involved in minor impacts, such as fender-benders, or where the nature of the crash—such as a collision involving parked cars—does not subject the occupants to the full force of a high-speed crash.
+
+## Role of High Variation {#sec-discussion-variation}
+
+While the results of the analysis provide valuable insights into the factors influencing MVC fatalities, they also highlight significant uncertainties, particularly with some predictors such as the interaction terms.
+In @fig-coefficient-estimates, the variability of the interaction terms underscores the inherent challenges of modeling rare events such as fatalities, where small changes in conditions can lead to disproportionately large impacts.
+The uncertainty observed in these predictors is not just a statistical artifact but reflects the complexity and stochastic nature of accidents.
+These results highlight the need for caution when interpreting model outputs and point to the importance of further data collection, refinement, and more sophisticated modeling techniques to enhance the accuracy and reliability of predictions.
+
+The variability in the coefficients for predictors l suggests that there may be additional, unobserved factors influencing the likelihood of fatalities that are not captured by the current model.
+This uncertainty could stem from gaps in the available data or from the limitations in the types of variables collected, such as the lack of detailed road conditions, driver behavior, or vehicle-specific characteristics (e.g., model, age of the vehicle).
+To reduce this uncertainty, a more comprehensive data collection strategy is needed, focusing on variables that could provide a deeper understanding of accident dynamics, such as weather conditions, road infrastructure quality, or even the socio-economic status of the areas where accidents occur.
+By expanding the scope and granularity of the data, it would be possible to refine the model, improving the precision of the estimates and providing clearer insights into the factors that contribute to fatalities.
+
+The variability in the model’s predictions also speaks to the stochastic nature of accidents, where the outcome of any given crash is influenced by a multitude of unpredictable factors.
+MVC fatalities, while significant, are relatively rare compared to non-fatal accidents, making them inherently difficult to predict.
+In this context, the uncertainty observed in the model's results can be understood as a reflection of the randomness and unpredictability involved in traffic collisions.
+Factors such as driver response times, unexpected road hazards, or even the randomness of the sequence of events leading up to a fatal crash are beyond the scope of even the most sophisticated predictive models.
+In addition, fatalities often result from a confluence of circumstances—such as high speed, poor weather conditions, or the failure of safety mechanisms—that may not be fully captured by existing data or model specifications.
+
+## Implications {#sec-discussion-implications}
+
+The findings from this analysis provide valuable insights for policymakers and law enforcement agencies seeking to improve road safety.
+One of the most pressing implications is the identification of high-risk periods, particularly during rush hours and late evenings, where the likelihood of fatal collisions is significantly higher.
+Targeted interventions could focus on enforcing speed limits and impaired driving laws during these critical hours, which may help mitigate the increased risk of accidents.
+For example, police could increase patrols and sobriety checkpoints, particularly in areas with higher concentrations of motorcycle or pedestrian collisions.
+Moreover, understanding the time of day when accidents are most likely to result in fatalities allows for more efficient deployment of resources, ensuring that law enforcement is present during times when their intervention could have the greatest impact on reducing fatalities.
+
+In addition to targeted law enforcement, the analysis also points to infrastructure improvements that could reduce fatal outcomes for vulnerable groups, such as pedestrians and motorcyclists.
+Police and local governments should prioritize the creation of safer urban environments by enhancing lighting in poorly lit areas and developing separated lanes for bicycles and motorcycles.
+These measures could help reduce the exposure of vulnerable road users to traffic hazards, particularly during high-risk times.
+Furthermore, predictive insights from statistical modeling could revolutionize real-time road safety strategies.
+For instance, traffic management systems could use predictive data to adjust traffic light patterns during peak hours, optimizing traffic flow and reducing congestion-related accidents.
+Additionally, emergency services could use data-driven insights to allocate resources more efficiently, ensuring that areas with higher likelihoods of fatal collisions are prioritized for immediate response.
+In this way, statistical modeling can provide law enforcement and urban planners with the tools necessary to not only react to but proactively address high-risk situations on the road.
+
+## Weaknesses and Next Steps {#sec-discussion-weaknesses}
+
+The model’s limitations stem from both the stochastic nature of traffic accidents and potential biases in the data.
+Fatalities are rare events, and the randomness inherent in accident outcomes—such as variations in road conditions, driver behavior, or other small factors—makes it difficult for models to predict them with high precision.
+Additionally, biases in data collection, like inconsistent reporting of certain variables or underreporting of less severe accidents, may lead to seleciton bias towards more severe accidnets in our datasets, affecting our predictions.
+For example, in the United States, almost a third of car crashes involving a vulnerable road user go unreported (INSERT CITATION).
+Though this is a statistics from the United States, the article's findings highlight that the potential biases present in our data.
+These challenges introduce uncertainties in the model’s estimates, particularly for less predictable outcomes like fatalities.
+
+Moreover, the model’s reliance on aggregate data limits its ability to account for the nuanced, dynamic nature of real-time road events.
+The factors influencing a fatality may not be static and can vary significantly based on the specific context of each collision.
+This variability makes it challenging to create a one-size-fits-all model for MVC fatalities, and the results are likely to change with shifts in environmental, societal, or technological factors.
+To improve accuracy and robustness, further data collection is essential, particularly to capture rare but impactful events and refine predictors of fatality outcomes.
+Enhancements in data quality and the incorporation of more granular, real-time data will improve model predictions and provide better insight into the complex dynamics of traffic accidents.
+
+As a next step, future research should focus on collecting more granular, real-time data to capture the dynamic factors influencing MVC fatalities.
+This could involve leveraging emerging technologies such as smart traffic sensors, vehicle telematics, and data from wearable devices to capture more detailed and context-specific information.
+Additionally, refining statistical techniques, such as Bayesian hierarchical models or machine learning approaches, can help better account for the inherent randomness of accidents and the uncertainty in predicting rare outcomes.
+Improved modeling could lead to more accurate, actionable insights for policy development and real-time road safety interventions.
 
@@
 
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In we implement a posterior predictive check. This shows...
-
-In we compare the posterior with the prior. This shows...
-
-```{r}
+## Exploration of Surveys, Sampling, and Observational Data
+
+### Data Collection Methodology
+
+The Toronto Police Service's Motor Vehicle Collision database is collected using a systematic process.
+When a collision occurs such that where persons are injured or a combined damage is valued at more than $\2000 to vehicles or property, or damage to any private, municipal or highway property, police officers complete a standardized Motor Vehicle Accident Report form.
+This data is then maintained by the City of Toronto, where the collection process follow a standardized process highlighted in the Motor Vehicle Collision Report manual (CITATION), ensuring reporting consistency.
+The dataset also defines a collision: the contact resulting from the motion of a motor vehicle or a street car of its load, which produces property damage, injury or death.
+A standardized definition ensures consistent reporting by the Toronto Police Service and highlights the dataset's clear boundaries.
+
+For every accident that requires one to fill out the Motor Vehicle Accident Report form, the police officer fills out the SR-LD-402 form (CITATION).
+The process ensures consistency in reporting key details surrounding a motor vehicle accident.However, because police officers only need to fill out this form in specific instances, the data set only includes collisions where police officers were present or serious incidents that required a report.
+As a result, minor offences are usually left out in the dataset and often go unreported.
+
+The dataset primarily focuses on all types of Motor Vehicle Collisions.
+The dataset specifically focuses on property damage collisions, fail to remain collisions, injury collisions and fatalities.
+Fatal Collisions occur when an individual’s injuries from a MVC result in a fatality within 30 days.
+Fatal Collisions excludes occurrences on private property, ones related to sudden death prior to collision, such as suicide, and occurrences where the individual has died more than 30 days after the collision.
+Personal Injury Collisions occur when an individual involved in a MVC suffers personal injuries.
+Fail to Remain Collisions occur when an individual involved in a MVC fails to stop and provide their information at the scene of a collision.
+Property Damage Collisions occur when an individual’s property has been damaged in a MVC or the value of damages is less than $2,000 for all involved These defintions, specifically the one for Fatal Collisions, sets a clear boundary for data inclusion, but excludes motor vehicle accidents that occur on private property.
+
+Each entry in the dataset represents a single Motor Vehicle Collision incident.
+Therefore, multiple offences or victims can be associated with each record.
+This means that a singular person might have multiple records if they have gotten into multiple Motor Vehicle Collision incidents.
+
+A notable aspect of the dataset is its handling of spatial data.
+Each collision includes the approximate latitude and longitude coordinates.
+The location of calls for service have been deliberately offset to the nearest road intersection node to protect the privacy of parties.
+Because of the offset of offences location, the coordinates may not reflect the exact count of occurrences.The statistics at the neighborhood level may not reflect the true location of the collisions.
+The Toronto Police Service acknowledges these limitations and does not guarantee the accuracy, completeness, timeliness of the data, and TPS encourages to not compare it to any source of crime data.
+
+### Idealized Methodology
+
+The current data collection methodology employed by the Toronto Police Service provides valuable insights into Motor Vehicle Collisions (MVCs), yet several enhancements could make the dataset more comprehensive and actionable.
+Expanding the criteria for reporting is a critical first step.
+By including minor collisions that do not meet the $2,000 damage threshold or require police presence, the city can gain a fuller understanding of road safety dynamics.
+A self-reporting mechanism, such as an online portal or app, would allow drivers and witnesses to log details of these incidents conveniently.
+Additionally, redefining fatality criteria to include deaths occurring beyond 30 days due to collision-related injuries would align the dataset with public health standards, ensuring fatalities are not undercounted.
+
+Incorporating advanced data collection techniques could significantly enhance data accuracy and granularity.
+Real-time telemetry from vehicles, collected via partnerships with automakers and insurers, can provide precise information on collision dynamics such as speed, braking patterns, and collision forces.
+Crowdsourcing data through a city-wide mobile app could also capture unreported incidents and near-misses, enriching the dataset with granular environmental and behavioral details.
+Integration of traffic camera data, analyzed using AI-powered algorithms, would add an unbiased and continuous data stream to validate collision reports and identify patterns.
+These methods collectively address gaps in the current system, providing deeper insights into the factors contributing to MVCs.
+
+Expanding the scope of variables collected in collision reports is another avenue for improvement.
+Adding details about driver behavior (e.g., speeding or distracted driving), the involvement of emerging vehicle technologies (e.g., self-driving systems), and the road environment (e.g., construction zones) can offer richer insights into the conditions that contribute to MVCs.
+This would help policymakers and researchers better understand the interplay of human, environmental, and technological factors in collision outcomes.
+
+### Example Survey
+
+To gather more information about motor vehicle collisions, I've attached a proposed survey to better understand the factors that contribute to fatalities and injuries.
+We aim to replicate something similar to the Motor Vehicle Collision Report and aim to collect data for car crashes in Toronto.
+All responses will be confidential.
+I've attached the proposed survey [here](https://forms.gle/C1uBXbbUddNvZMR4A).
+
+\newpage
+
+## Enhancement: Model Card
+
+### Model Overview
+This model predicts the occurrence of motor vehicle collisions (MVCs) based on various factors, including the time of day (hour), type of vehicle involved (motorcycle, pedestrian, automobile, etc.), and police division.
+The model is trained using data from 2015 to 2024, with a focus on the interactions between the hour of the day and motorcycle or pedestrian involvement.
+
+### Model Details
+
+- **Model Type**: Bernoulli regression (logistic regression for binary outcomes)
+- **Target Variable**: Binary classification of motor vehicle collisions (occurrence of MVCs)
+- **Training Data**: 1606 observations
+- **Features**: 
+  - Hour of the day (`hour1` to `hour23` representing 24 hours in a day)
+  - Vehicle type (motorcycle, pedestrian, automobile, etc.)
+  - Type of accident (injury_collision, fail_to_remain_collision, and property_damage_collision)
+  - Police division (D12, D13, etc.)
+  - Year of the accident (2015–2024)
+
+### Model Coefficients
+
+- **Intercept**: 9.705
+- **Hour Coefficients**:
+  - `hour1` through `hour23`: Varies with each hour of the day. Positive coefficients for hours between 2 PM to 11 PM suggest higher accident probability in these hours.
+  - **Key Hours**:
+    - `hour2`: 1.920
+    - `hour3`: 2.908
+    - `hour6`: 0.966
+    - `hour7`: -1.673
+    - `hour8`: -0.511
+    - `hour21`: 1.237
+    - `hour22`: 0.631
+    - `hour23`: 1.572
+- **Vehicle Type Coefficients**:
+  - **Motorcycle**: 17.036, indicating a high likelihood of motorcycle-related collisions.
+  - **Pedestrian**: 6.832, showing the significance of pedestrian involvement in accidents.
+  - **Bicycle**: 3.809, showing a moderate likelihood of bicycle-related accidents.
+  - **Automobile**: 2.648, indicating that automobile-related collisions are also common.
+- **Collision Type Coefficients**:
+  - **Injury Collision**: -34.794, indicating a strong negative association with injury-related collisions.
+  - **Fail to Remain Collision**: -29.912, indicating a strong negative association with collisions where the driver fails to remain.
+  - **Property Damage Collision**: -30.004, showing a significant negative impact for property damage-related collisions.
+- **Police Division Coefficients**:
+  - `police_divisionD12`: 0.846, indicating division-specific effects, where division D12 has a positive association with collisions.
+  - `police_divisionD13`: 0.075
+  - `police_divisionD14`: -2.806
+  - `police_divisionD22`: 1.797
+  - `police_divisionD23`: 1.193
+  - `police_divisionD31`: 0.619
+  - `police_divisionD32`: 1.007
+  - `police_divisionD33`: 0.055
+  - `police_divisionD41`: 0.989
+  - `police_divisionD42`: 1.140
+  - `police_divisionD43`: 0.582
+  - `police_divisionD51`: 0.915
+  - `police_divisionD52`: 1.465
+  - `police_divisionD53`: 1.389
+  - `police_divisionD55`: 0.389
+  - `police_divisionNSA`: -4.040
+- **Year Coefficients**:
+  - `year2015`: -0.500
+  - `year2016`: 0.329
+  - `year2017`: 0.412
+  - `year2018`: 0.165
+  - `year2019`: -0.025
+  - `year2020`: 0.021
+  - `year2021`: 0.079
+  - `year2022`: 1.947, showing a significant increase in collisions during this year.
+  - `year2023`: -1.708, showing a decrease in collisions during this year.
+  - `year2024`: -0.210
+- **Interaction Terms**:
+  - **Hour × Motorcycle**:
+    - `hour1 × motorcycle`: 31.900
+    - `hour2 × motorcycle`: 33.094
+    - `hour3 × motorcycle`: 33.377
+    - `hour4 × motorcycle`: 33.633
+    - `hour6 × motorcycle`: 51.053
+    - `hour7 × motorcycle`: 52.055
+    - `hour8 × motorcycle`: 55.746
+    - `hour9 × motorcycle`: -66.051
+    - `hour10 × motorcycle`: 27.802
+    - `hour12 × motorcycle`: 51.049
+    - `hour13 × motorcycle`: 18.620
+    - `hour14 × motorcycle`: -12.621
+    - `hour15 × motorcycle`: -11.489
+    - `hour16 × motorcycle`: 32.651
+    - `hour17 × motorcycle`: 20.150
+    - `hour18 × motorcycle`: 12.044
+    - `hour19 × motorcycle`: -11.899
+    - `hour20 × motorcycle`: -5.807
+    - `hour21 × motorcycle`: 21.915
+    - `hour22 × motorcycle`: 9.369
+    - `hour23 × motorcycle`: 34.897
+  - **Hour × Pedestrian**:
+    - `hour1 × pedestrian`: 15.095
+    - `hour2 × pedestrian`: 29.594
+    - `hour3 × pedestrian`: 20.771
+    - `hour4 × pedestrian`: 19.986
+    - `hour5 × pedestrian`: 23.943
+    - `hour6 × pedestrian`: 9.778
+    - `hour7 × pedestrian`: 1.434
+    - `hour8 × pedestrian`: 0.484
+    - `hour9 × pedestrian`: 1.831
+    - `hour10 × pedestrian`: 1.895
+    - `hour11 × pedestrian`: 15.280
+    - `hour12 × pedestrian`: 9.497
+    - `hour13 × pedestrian`: 1.397
+    - `hour14 × pedestrian`: 12.171
+    - `hour15 × pedestrian`: -0.552
+    - `hour16 × pedestrian`: 0.453
+    - `hour17 × pedestrian`: 0.272
+    - `hour18 × pedestrian`: -5.210
+    - `hour19 × pedestrian`: -0.707
+    - `hour20 × pedestrian`: 7.720
+    - `hour21 × pedestrian`: 12.510
+    - `hour22 × pedestrian`: 1.760
+    - `hour23 × pedestrian`: -2.690
+
+#### Interaction Terms
+
+- **Hour and Motorcycle**:
+  - `hour1 × motorcycle`: 31.900, significant interaction in early morning hours, suggesting motorcycle accidents are more likely during certain hours.
+  - Other significant hours for motorcycle collisions are `hour6 × motorcycle` (51.053) and `hour9 × motorcycle` (-66.051).
+- **Hour and Pedestrian**:
+  - `hour1 × pedestrian`: 15.095, significant pedestrian accidents in the early morning hours.
+  - Other significant hours include `hour2 × pedestrian` (29.594) and `hour5 × pedestrian` (23.943), highlighting peak pedestrian accident hours.
+
+### Model Performance Metrics
+
+- **R²**: 0.992, indicating excellent fit to the data.
+- **Log Likelihood**: -0.082
+- **Effective Log Predictive Density (ELPD)**: -16.2, with an associated standard error of 2.2.
+- **Leave-One-Out Information Criterion (LOOIC)**: 32.4, with a standard error of 4.4.
+- **Watanabe-Akaike Information Criterion (WAIC)**: 21.3
+- **Root Mean Squared Error (RMSE)**: 0.00, indicating near-perfect predictive accuracy.
+
+### Training Time
+- **Chain 1**:
+  - Warm-up: 29.99 seconds
+  - Sampling: 26.11 seconds
+  - Total: 56.1 seconds
+- **Chain 2**:
+  - Warm-up: 19.62 seconds
+  - Sampling: 17.23 seconds
+  - Total: 36.84 seconds
+- **Chain 3**:
+  - Warm-up: 19.88 seconds
+  - Sampling: 19.30 seconds
+  - Total: 39.18 seconds
+- **Chain 4**:
+  - Warm-up: 22.40 seconds
+  - Sampling: 19.17 seconds
+  - Total: 41.57 seconds
+
+### Conclusion
+The model performs exceptionally well in predicting motor vehicle collisions based on hour, vehicle type, and other factors, with the highest predictive power for motorcycle-related accidents.
+Key interactions, such as between hour and motorcycle or pedestrian, highlight the importance of specific hours of the day in predicting collision occurrences.
+This model can be used for more effective traffic safety planning and intervention strategies.
+
+\newpage
+
+## Data Attribution Statement {#sec-appendix-attribution}
+
+"This data contains information licensed under the Open Government License - Toronto" [@tphlicense].
+
+## Model Diagnostics {#sec-appendix-diagnostics}
+
+@fig-ppcheck compares observed data (dark line) with replicated posterior predictions (light blue lines).
+The close alignment suggests that the model accurately captures the data's central tendency and variability.
+@fig-convergencecheckTrace and @fig-convergencecheckRhat show that the sampling algorithm used, the Markov chain Monte Carlo (MCMC) algorithm, did not run into issues as the posterior distribution for the model was created.
+Using the checks presented by @citetellingstorieswithdata, both graphs do not show anything abnormal since the trace plots in @fig-convergencecheckTrace display substantial horizontal fluctuation across chains, indicating good mixing, while the Rhat values in @fig-convergencecheckRhat are close to 1 and well below 1.05, further supporting convergence.
+
+```{r}
+#| echo: false
 #| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-#pp_check(first_model) +
-  #theme_classic() +
- # theme(legend.position = "bottom")
-
-#posterior_vs_prior(first_model) +
- # theme_minimal() +
-  #scale_color_brewer(palette = "Set1") +
- # theme(legend.position = "bottom") +
- # coord_flip()
-```
-
-## Diagnostics
-
-is a trace plot. It shows... This suggests...
-
-is a Rhat plot. It shows... This suggests...
+#| message: false
+#| warning: false
+#| label: fig-ppcheck
+#| fig-cap: "Posterior Predictive Check: Comparison of Observed and Replicated Data"
+pp_check(prediction_model) + theme_minimal()
+```
 
@@
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-#plot(first_model, "trace")
-
-#plot(first_model, "rhat")
+#| label: fig-convergencecheckTrace
+#| fig-cap: "Checking the convergence of the MCMC algorithm - Trace"
+trace_plot <- plot(prediction_model, "trace") + theme_minimal()
+grid.arrange(trace_plot, ncol = 5, nrow = 5)
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-convergencecheckRhat
+#| fig-cap: "Checking the convergence of the MCMC algorithm - Rhat"
+plot(prediction_model, "rhat") + theme_minimal()
 ```

### 2024-12-03 paper/paper.qmd
added title and abstract

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Predicting Fatalities in Toronto Car Crashes: Analyzing the Impact of Time, Vehicle Type, and Other Collision Factors"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study aims to predict the likelihood of fatality in motor vehicle collisions (MVCs) in Toronto from 2014 to 2023, using data on crash time, vehicle type, and collision characteristics.
+By analyzing the interaction between hour of the day and vehicle types, including motorcycles and pedestrians, the model identifies key factors contributing to fatal outcomes.
+Results indicate that late-afternoon to evening hours, particularly between 4 PM and 11 PM, show a higher likelihood of fatalities, with motorcycles and pedestrians significantly increasing fatality risk.
+The model achieved a high predictive accuracy, demonstrating its potential in informing safety policies and interventions in Toronto."
 format: pdf
@@
 ggplot(pedestrian_fatalities, aes(x = hour, y = pedestrian_fatalities)) +
-  geom_point(color = "red", size = 2) +
+  geom_line(color = "green", size = 2) +
   theme_minimal() +

### 2024-12-03 paper/paper.qmd
added citations and updated intext referencing

@@
 
-The estimand of interest is the log-adjusted probability of a fatality occuring in a MVC.
-Specifically, we aim to quanitfy how specific environmental factors increase or decrease the likelihood of a fatality.
-By applying inferential analysis through Bayesian linear models, we assess not only the magnitude of these effects, but their underlying uncertainties (EDIT).
+The estimand of interest is the log-adjusted probability of a fatality occurring in a MVC.
+Specifically, we aim to quantify how specific environmental factors increase or decrease the likelihood of a fatality.
+By applying inferential analysis through Bayesian linear models, we assess not only the magnitude of these effects, but their underlying uncertainties.
 
@@
 The remainder of this paper is structured as follows: @sec-data describes the dataset and methodology and @sec-model exhibits the use of inferential models.
-@Sec-results presents the results of the analysis, detailing the observed relationships between likelihood of death and various circumstantial factors.
+@sec-results presents the results of the analysis, detailing the observed relationships between likelihood of death and various circumstantial factors.
 @sec-discussion discusses the broader implications and limitations of our findings.@sec-appendix presents a detailed idealized methodology to improve data collection, and additional model summary and diagnostic information.
@@
 
-## Overview
-
-This dataset, "Police Annual Statistical Report - Traffic Collisions", was published and refreshed on October 21st, 2024, by the Toronto Police Service [insert citation].
+## Overview {#sec-data-overview}
+
+This dataset, "Police Annual Statistical Report - Traffic Collisions", was published and refreshed on October 21st, 2024, by the Toronto Police Service [@vehiclecollision].
 The Toronto Police Service publishes various datasets on public safety and crime to inform the public about safety issues [@annual_statistics_report].
@@
 
-There is an alternative dataset from the Toronto Police Service called "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" (CITE).
+There is an alternative dataset from the Toronto Police Service called "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" [@alternativedataset].
 Unlike the alternative dataset, this paper's dataset focuses on all collisions, instead of only focusing on ones where someone was either killed or seriously injured.
@@
 The paper uses the R programming language [@citeR] to analyze the dataset.
-The tidyverse package was used to simulate the dataset.
-Also, the tidyverse [@citetidyverse], arrow [CITE] and opendatatoronto [@citeopendatatoronto] packages were used to download the Victims of Crime dataset.
-Then, the tidyverse [@citetidyverse] package was used to clean the raw dataset and generate tests.
-The testthat package [CITE THIS] was used to create tests for our cleaned dataset.
-Rstanarm [CITE], Arrow [CITE], and bayestestR [CITE] were used to create and test the model.
-Finally, ggplot2 [@citeggplot2], tidyverse [@citetidyverse], knitr [@citeknitr] and scales [@citescales] packages were used to create the tables and graphs to display the results.
+The tidyverse package was used to simulate the dataset [@tidyverse].
+Also, the tidyverse [@tidyverse], here [@here], arrow [@citearrow] and opendatatoronto [@citeopendatatoronto] packages were used to download the Victims of Crime dataset.
+Then, the tidyverse [@tidyverse] package was used to clean the raw dataset and generate tests.
+The testthat package [@citetestthat] was used to create tests for our cleaned dataset.
+rstanarm [@rstanarm], Arrow [@citearrow], and bayestestR [@citebayestestr] were used to create and test the model.
+Finally, ggplot2 [@ggplot2], tidyverse [@tidyverse], knitr [@citeknitr] and scales [@citescales] packages were used to create the tables and graphs to display the results.
 [edit this paragraph]
 
-## Measurement
+## Measurement {#sec-data-measurement}
 
 Transforming a real-life Motor Vehicle Collision to an entry in the dataset is a well-documented process by the Toronto Police Service.
-For insurance purposes, the Toronto Police Service requires drivers to fill out the Motor Vehicle Collision Report for any collisions that occur in Toronto (CITE).
-Drivers required to fill out a motor vehicle collision report if the combined damage is more than $2000, if someone is injured, if a criminal act such as a DUI occurs, or if a pedestrian is involved in the accident (CITE).
+For insurance purposes, the Toronto Police Service requires drivers to fill out the Motor Vehicle Collision Report for any collisions that occur in Toronto [@tps_mvc_reports].
+Drivers required to fill out a motor vehicle collision report if the combined damage is more than $2000, if someone is injured, if a criminal act such as a DUI occurs, or if a pedestrian is involved in the accident [@tps_mvc_reports].
 These reports are retained for six years by the Toronto Police Service, with the exception of collisions resulting in a fatality, which are retained indefinitely.
@@
 
-## Outcome variables
+## Outcome variables {#sec-data-outcome}
 
@@
 ```
-However, after the transformation, the outcome variable now takes on binary values and the distribution and summary statistics are shown below in @fig-fatalites-cleaned and @tbl-fatalites:
+However, after the transformation, the outcome variable now takes on binary values and the distribution and summary statistics are shown below in @fig-fatalities-cleaned and @tbl-fatalities:
 
@@
 ```
-## Predictor variables
-
-### Hour
+## Predictor variables {#sec-data-predictor}
+
+### Hour {#sec-data-predictor-hour}
 
@@
 
-### Type of Crash
+### Type of Crash {#sec-data-predictor-crash}
 
@@
 
-### Automobile
+### Automobile {#sec-data-predictor-automobile}
 
@@
 
-### Motorcycle
+### Motorcycle {#sec-data-predictor-motorcycle}
 
@@
 ```
-### Passenger
+### Passenger {#sec-data-predictor-passenger}
 
@@
 ```
-### Pedestrian
+### Pedestrian {#sec-data-predictor-pedestrian}
 
@@
 ```
-### Police Division
+### Police Division {#sec-data-predictor-police}
 
@@
 
-### Year
+### Year {#sec-data-predictor-year}
 
@@
 
-### Interaction Terms
+### Interaction Terms {#sec-data-predictor-interaction}
 
@@
 
-#### Hour and Motorcycle
+#### Hour and Motorcycle {#sec-data-predictor-interaction-hourmotorcycle}
 
@@
 
-#### Hour and Pedestrian
+#### Hour and Pedestrian {#sec-data-predictor-interaction-hourpedestrian}
 
@@
 ```
-# Model
-
-## Model Set-Up and Justification
+# Model {#sec-model}
+
+## Model Set-Up and Justification {#sec-model-setupjustification}
 
@@
 
-## Model Assumptions and Limitations
+## Model Assumptions and Limitations {#sec-model-assumptions}
 
@@
 
-# Results
+# Results {#sec-results}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-# Appendix {-}
-
-## Exploration of Surveys, Sampling, and Observational Data
-
-### Data Collection Methodology
+# Appendix {#sec-appendix}
+
+## Exploration of Surveys, Sampling, and Observational Data {#sec-appendix-sso}
+
+### Data Collection Methodology {#sec-appendix-sso-methodology}
 
@@
 When a collision occurs such that where persons are injured or a combined damage is valued at more than $\2000 to vehicles or property, or damage to any private, municipal or highway property, police officers complete a standardized Motor Vehicle Accident Report form.
-This data is then maintained by the City of Toronto, where the collection process follow a standardized process highlighted in the Motor Vehicle Collision Report manual (CITATION), ensuring reporting consistency.
+This data is then maintained by the City of Toronto, where the collection process follow a standardized process highlighted in the Motor Vehicle Collision Report manual [@mvcr_guide], ensuring reporting consistency.
 The dataset also defines a collision: the contact resulting from the motion of a motor vehicle or a street car of its load, which produces property damage, injury or death.
@@
 
-For every accident that requires one to fill out the Motor Vehicle Accident Report form, the police officer fills out the SR-LD-402 form (CITATION).
+For every accident that requires one to fill out the Motor Vehicle Accident Report form, the police officer fills out the SR-LD-402 form [@mvcr_guide].
 The process ensures consistency in reporting key details surrounding a motor vehicle accident.However, because police officers only need to fill out this form in specific instances, the data set only includes collisions where police officers were present or serious incidents that required a report.
@@
 
-### Idealized Methodology
+### Idealized Methodology {#sec-appendix-sso-idealized}
 
@@
 
-### Example Survey
+### Example Survey {#sec-appendix-sso-survey}
 
@@
 
-## Enhancement: Model Card
-
-### Model Overview
+## Model Card {#sec-appendix-modelcard}
+
+### Model Overview {#sec-appendix-modelcard-overview}
 This model predicts the occurrence of motor vehicle collisions (MVCs) based on various factors, including the time of day (hour), type of vehicle involved (motorcycle, pedestrian, automobile, etc.), and police division.
@@
 
-### Model Details
+### Model Details {#sec-appendix-modelcard-details}
 
@@
 
-### Model Coefficients
+### Model Coefficients {#sec-appendix-modelcard-coefficients}
 
@@
 
-#### Interaction Terms
+#### Interaction Terms {#sec-appendix-modelcard-interaction}
 
@@
 
-### Model Performance Metrics
+### Model Performance Metrics {#sec-appendix-modelcard-performance}
 
@@
 
-### Training Time
+### Training Time {#sec-appendix-modelcard-time}
 - **Chain 1**:
@@
 
-### Conclusion
+### Conclusion {#sec-appendix-modelcard-conclusion}
 The model performs exceptionally well in predicting motor vehicle collisions based on hour, vehicle type, and other factors, with the highest predictive power for motorcycle-related accidents.
@@
 @fig-convergencecheckTrace and @fig-convergencecheckRhat show that the sampling algorithm used, the Markov chain Monte Carlo (MCMC) algorithm, did not run into issues as the posterior distribution for the model was created.
-Using the checks presented by @citetellingstorieswithdata, both graphs do not show anything abnormal since the trace plots in @fig-convergencecheckTrace display substantial horizontal fluctuation across chains, indicating good mixing, while the Rhat values in @fig-convergencecheckRhat are close to 1 and well below 1.05, further supporting convergence.
+Using the checks presented by @tellingstories, both graphs do not show anything abnormal since the trace plots in @fig-convergencecheckTrace display substantial horizontal fluctuation across chains, indicating good mixing, while the Rhat values in @fig-convergencecheckRhat are close to 1 and well below 1.05, further supporting convergence.
 

### 2024-12-03 paper/paper.qmd
updated paper's readability and prose

@@
 ---
-title: "Predicting Fatalities in Toronto Car Crashes: Analyzing the Impact of Time, Vehicle Type, and Other Collision Factors"
+title: "Predicting Fatalities in Toronto Car Crashes: Analyzing the Impact of Time, Vehicle Type, and Other Collision Determinants"
 author: 
@@
 abstract: "This study aims to predict the likelihood of fatality in motor vehicle collisions (MVCs) in Toronto from 2014 to 2023, using data on crash time, vehicle type, and collision characteristics.
-By analyzing the interaction between hour of the day and vehicle types, including motorcycles and pedestrians, the model identifies key factors contributing to fatal outcomes.
+By analyzing the interaction between hour of the day and vehicle types, including motorcycles and pedestrians, the model identifies key elements contributing to fatal outcomes.
 Results indicate that late-afternoon to evening hours, particularly between 4 PM and 11 PM, show a higher likelihood of fatalities, with motorcycles and pedestrians significantly increasing fatality risk.
@@
 library(broom.mixed)
+library(scales)
 library(gridExtra)
@@
 
+\newpage
+
 # Introduction
@@
 Automobile fatalities are a leading cause of death worldwide, posing a public health and safety concern for urban cities.
-For example, in 2024, thirty people have died on Toronto's roadways so far, which is a 20% increase from last year (insert citation, CBC).
-However, new headlines, such as CBCs, highlight that Motor Vehicle Collisions (MVC) tend to either be generalized as a summary statistic or typically fatal events are over analyzed to the extent that environmental factors surrounding the crash are ignored (CLARIFY).
-Moreover, general environmental factors such as the time of the crash or if a bicycle was involved in the accident are important to understand what increases the likelihood of a fatality occurring in an MVC.
-The use of statistical modeling on increasingly available vehicle collision data presents an opportunity to develop a nuanced understanding of what factors increases the likelihood for a fatality to occur in an accident.
-This paper uses the Toronto Police Service's  Annual Statistical Report from Open Data Toronto to analyze what factors are most responsible in predicting if a fatality occurs in a MVC.
+For example, in 2024, thirty people have died on Toronto's roadways so far, which is a 20% increase from last year [@cbc_2024].
+However, news headlines, such as CBCs, highlight that Motor Vehicle Collisions (MVC) tend to either be generalized as a summary statistic or typically fatal events are over analyzed to the extent that environmental elements surrounding the crash are ignored.
+General environmental elements such as the time of the crash or if a bicycle was involved in the accident are important to understand what increases the likelihood of a fatality occurring in an MVC.
+The use of statistical modeling on increasingly available vehicle collision data presents an opportunity to develop a nuanced understanding of what variables increases the likelihood for a fatality to occur in an accident.
+This paper uses the Toronto Police Service's  Annual Statistical Report from Open Data Toronto to analyze what determinants are most responsible in predicting if a fatality occurs in a MVC.
 
 The estimand of interest is the log-adjusted probability of a fatality occurring in a MVC.
-Specifically, we aim to quantify how specific environmental factors increase or decrease the likelihood of a fatality.
+Specifically, we aim to quantify how specific environmental variables increase or decrease the likelihood of a fatality.
 By applying inferential analysis through Bayesian linear models, we assess not only the magnitude of these effects, but their underlying uncertainties.
 
-Our analysis identifies key predictors of vehicular accident fatalities, emphasizing heightened risks for motorcyclists (coefficient: 17.036) and pedestrians (coefficient: 6.809), especially during rush hours (hours 16 and 17) and late evenings (11:00 PM), where motorcycle collisions show a fatality likelihood of around 66%.
+Our analysis identifies key predictors of vehicular accident fatalities, emphasizing heightened risks for motorcyclists and pedestrians, especially during rush hours (4PM to 5PM) and late evenings (11:00 PM).
+Motorcycle collisions in the evening show a fatality likelihood of around 66%.
 Collision type variables, such as injury and property damage collisions, significantly lower fatality risks, while police division and year have minimal impact.
-However, variability in estimates, particularly for motorcyclist and automobile predictors, highlights uncertainty in their precise effects, underscoring the need for targeted interventions during high-risk times and for vulnerable populations.
-
-The paper is not only important from a public health perspective, but the paper also has policy development implications.
-Road safety and reducing fatal MVCs are a critical agenda item of any municipal government.
-The paper informs what factors increase the likelihood of death in an MVC, which informs policymakers' focus for relevant policy design.
+However, variability in estimates, particularly for motorcyclist and automobile predictors, highlights uncertainty in their precise effects, underscoring the need for targeted interventions during high-risk times and for vulnerable populations, informing policy decisions.
 
 The remainder of this paper is structured as follows: @sec-data describes the dataset and methodology and @sec-model exhibits the use of inferential models.
-@sec-results presents the results of the analysis, detailing the observed relationships between likelihood of death and various circumstantial factors.
-@sec-discussion discusses the broader implications and limitations of our findings.@sec-appendix presents a detailed idealized methodology to improve data collection, and additional model summary and diagnostic information.
+@sec-results presents the results of the analysis, detailing the observed relationships between likelihood of death and various circumstantial variables.
+@sec-discussion discusses the broader implications and limitations of our findings.
+@sec-appendix presents a detailed idealized methodology to improve data collection, a model card, and additional model summary and diagnostic information.
 
@@
 Data on traffic collisions is included in the Toronto Police Service's Annual Statistical Report, which also covers reported crimes, search of persons, firearms, and the Police Service's budget [@annual_statistics_report].
-The data is collected using historical Motor Vehicle Collisions and classifies them into the following categories:
-* Property Damage (PD) Collisions
-* Fail to Remain (FTR) Collisions, or commonly known as hit-and-run accidents
-* Injury Collisions
-* Fatalities
+The data is collected using historical Motor Vehicle Collisions and classifies them into the following categories: Property Damage (PD) Collisions; Fail to Remain (FTR) Collisions, or commonly known as hit-and-run accidents; Injury Collisions; and Fatalities
 
@@
 There is an alternative dataset from the Toronto Police Service called "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" [@alternativedataset].
-Unlike the alternative dataset, this paper's dataset focuses on all collisions, instead of only focusing on ones where someone was either killed or seriously injured.
-While the alternative dataset has more explanatory variables simply because more data is collected when someone dies or is seriously injured, this paper's aims to generalize if a fatality is more likely to occur based on the general circumstances surrounding a crash, such as the time of day or if property damage occurred.
+Though the alternative dataset has more explanatory variables due to the severity of the collisions recorded, this paper's selected dataset is more relevant for the focus of this paper.
+This dataset focuses on all types of Motor Vehicle Collisions, unlike the alternative dataset, which focuses on collisions where someone was killed or seriously injured.
 In addition, each entry in the alternative dataset represents a killed or injured person, meaning that there exists the possibility for duplicate entries.
-However, each entry in the selected dataset is a unique Motor Vehicle Collision, making it more suitable for the scope of this paper Thus, we ended up not going with the alternative dataset for this paper, but there are variables in the alternative dataset that may motivate future research on this subject.
-(EDIT TO MAKE MORE CLEAR)
+However, each entry in the selected dataset is a unique MVC, making it more suitable for the scope of this paper.
+Thus, we ended up not selecting the alternative dataset for this paper, but there are variables in the alternative dataset, such as alcohol use or road conditions, that may motivate future research on the subject.
 
@@
 rstanarm [@rstanarm], Arrow [@citearrow], and bayestestR [@citebayestestr] were used to create and test the model.
-Finally, ggplot2 [@ggplot2], tidyverse [@tidyverse], knitr [@citeknitr] and scales [@citescales] packages were used to create the tables and graphs to display the results.
-[edit this paragraph]
+Finally, ggplot2 [@ggplot2], tidyverse [@tidyverse], broom.mixed [@citebroommixed], bayesplot [@citebayesplot], sjPlot [@citesjplot], knitr [@citeknitr], and scales [@citescales] packages were used to create the tables and graphs to display the results.
 
@@
 For every collision, basic facts such as the location, time, and date of the collision is recorded through the Motor Vehicle Collision Report.
-Majority if not all the factors recorded in the dataset are all objective measurements regarding the specific details such as if a motorcycle was involved in the collision or if the collision resulted in property damage.
+Majority if not all the variables recorded in the dataset are all objective measurements regarding the specific details such as if a motorcycle was involved in the collision or if the collision resulted in property damage.
 All of these details are recorded in the Motor Vehicle Collision Report for all vehicular collisions and are entered into the dataset.
@@
 According to the dataset, a fatal collision occurs when an individual's injuries from a collision results in a fatality within 30 days.
-Fatal collisions excludes occurences on private property, ones related to suddend eath prior to collision, such as suicide, and where the individuals has died more than 30 days after the collision.
+Fatal collisions excludes occurrences on private property, ones related to sudden death prior to collision, such as suicide, and where the individuals has died more than 30 days after the collision.
 However, because we are more interested in predicting the probability that a fatality occurs than the number of fatalities, we transformed the variable that distinguishes collisions between if the collision resulted in any fatalities and those without fatalities.
@@
 
-# Plot distribution of Fatalies in the Raw Dataset 
+# Plot distribution of Fatalities in the Raw Dataset 
 ggplot(raw_data %>% mutate(FATALITIES = replace_na(FATALITIES, 0)), 
@@
        y = "Count") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  scale_y_continuous(labels = label_number(accuracy = 1))
 
@@
 #| label: tbl-fatalities
-#| tbl-cap: Number of unique high-quality polling organizations
+#| tbl-cap: Summary Statistics of the Fatalities Variable
 #| tbl-align: center
@@
 
+\newpage
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-hour
+#| tbl-cap: Summary of the Hour Variable 
+#| tbl-align: center
+#| tbl-width: 82%
+
+# Calculate summary statistics for the 'hour' variable
+summary_stats <- analysis_data %>%
+  group_by(hour) %>%
+  summarize(
+    "No Fatalities" = sum(fatalities == 0, na.rm = TRUE),
+    "Fatality Occurred" = sum(fatalities == 1, na.rm = TRUE),
+    Total = n()
+  ) %>%
+  rename(Hour = hour)
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable(col.names = c("Hour", "No Fatalities", "Fatality Occurred", "Total")) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+
 ```{r}
@@
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| label: tbl-hour
-#| tbl-cap: The Summary of the Hour 
-#| tbl-align: center
-#| tbl-width: 82%
-
-# Calculate summary statistics for the 'hour' variable
-summary_stats <- analysis_data %>%
-  group_by(hour) %>%
-  summarize(
-    "No Fatalities" = sum(fatalities == 0, na.rm = TRUE),
-    "Fatality Occurred" = sum(fatalities == 1, na.rm = TRUE),
-    Total = n()
-  )
-
-# Display the summary statistics in a nicely formatted table
-summary_stats %>%
-  kable() %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
-
-```
-
 ### Type of Crash {#sec-data-predictor-crash}
@@
 Fail to remain collisions are recorded when an individual involved in a collision fail to stop and provide their information at the scene of a collision.
-Property damage collisions occur when an individual has been damaged in a collision or the value of damages is less than $\2000 for all parties.
+Property damage collisions occur when an individual has been damaged in a collision or the value of damages is less than \$2000 for all parties.
 The distribution and summary statistics of these three variables can be found in @fig-crashtype and @tbl-crashtype.
@@
   scale_fill_manual(values = c("0" = "green", "1" = "red"), 
-                    labels = c("0" = "No Fatalities", "1" = "Fatality Occurred")) +
+                    labels = c("0" = "No Fatalities", "1" = "Fatality Occurred"),
+                    name = "Fatalities") +
   scale_x_discrete(labels = c("injury_collision" = "Injury Collision", 
                               "fail_to_remain_collision" = "Fail to Remain Collision", 
-                              "property_damage_collision" = "Property Damage Collision"))
+                              "property_damage_collision" = "Property Damage Collision")) +
+  scale_y_continuous(labels = label_number(scale = 1, accuracy = 1))
 
@@
 
-# Calculate summary statistics for the 'Collision Type' variable
+# Calculate summary statistics for the 'Collision Type' variables
 summary_stats <- analysis_data %>%
@@
 summary_stats %>%
-  kable() %>%
+  kable(col.names = c("Collision Type", "No Fatalities", "Fatality Occurred", "Total")) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
@@
 
-# Plot distribution of Hour Variable
+# Plot distribution of Automobile Variable
 ggplot(analysis_data, aes(x = automobile)) +
@@
        y = "Count") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-```
-
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  scale_y_continuous(labels = label_number(accuracy = 1))
+
+```
+\newpage
 ```{r}
@@
 
-# Calculate summary statistics for the 'hour' variable
+# Calculate summary statistics for the 'automobile' variable
 summary_stats <- analysis_data %>%
@@
 summary_stats %>%
-  kable() %>%
+  kable(col.names = c("Automobile", "No Fatalities", "Fatality Occurred", "Total")) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
@@
 
-# Calculate summary statistics for the 'hour' variable
+# Calculate summary statistics for the 'Motorcycle' variable
 summary_stats <- analysis_data %>%
@@
 summary_stats %>%
-  kable() %>%
+  kable(col.names = c("Hour", "No Fatalities", "Fatality Occurred", "Total")) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
 ```
+
 ### Passenger {#sec-data-predictor-passenger}
@@
 summary_stats %>%
-  kable() %>%
+  kable(col.names = c("Passenger", "No Fatalities", "Fatality Occurred", "Total")) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
@@
 @tbl-pedestrian highlights that of 606 deaths, 342 deaths involved pedestrians, which is a significant percentage.
-
-```{r}
-#| label: fig-pedestrian
-#| fig-cap: Distribution of the Pedestrian Variable  
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 5
-#| fig-height: 2.5
-#| fig-align: center
-
-# Plot distribution of Hour Variable
-ggplot(analysis_data, aes(x = pedestrian)) +
-  geom_bar(fill = "red", alpha = 0.7, color = "black") +
-  theme_minimal() +
-  labs(x = "Pedestrian",
-       y = "Count") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-```
 
@@
 summary_stats %>%
-  kable() %>%
+  kable(col.names = c("Pedestrian", "No Fatalities", "Fatality Occurred", "Total")) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
+```{r}
+#| label: fig-pedestrian
+#| fig-cap: Distribution of the Pedestrian Variable  
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Plot distribution of Hour Variable
+ggplot(analysis_data, aes(x = pedestrian)) +
+  geom_bar(fill = "red", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Pedestrian",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  scale_y_continuous(labels = label_number(accuracy = 1))
+
 ```
@@
 ```
+\newpage
 
@@
 summary_stats %>%
-  kable() %>%
+  kable(col.names = c("Police Division", "No Fatalities", "Fatality Occurred", "Total")) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
@@
 ```
-
+\newpage
 ```{r}
@@
 summary_stats %>%
-  kable() %>%
+  kable(col.names = c("Year", "No Fatalities", "Fatality Occurred", "Total")) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
@@
 @fig-hour-motorcycle shows that a significant number of deaths come in the evening from 2 PM to 11 PM.
-
-```{r}
-#| label: fig-hour-motorcycle
-#| fig-cap: Fatality Likelihood by Hour for Motorcyclists
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-width: 5
-#| fig-height: 2.5
-#| fig-align: center
-
-# Create interaction plot data for motorcycle
-interaction_motorcycle <- analysis_data %>%
-  filter(motorcycle == 1) %>%
-  group_by(hour) %>%
-  summarize(fatality_rate_motorcycle = mean(fatalities == 1, na.rm = TRUE))
-
-# Plot the interaction between hour and motorcycle
-ggplot(interaction_motorcycle, aes(x = hour, y = fatality_rate_motorcycle)) +
-  geom_line(color = "red", size = 1.2) +
-  labs(x = "Hour of Day", y = "Motorcycle Fatality Rate") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-```
 
@@
 #| tbl-align: center
+#| tbl-length: 50%
 #| tbl-width: 82%
@@
 summary_stats_motorcycle %>%
-  kable() %>%
+  kable(col.names = c("Hour", "Motorcyclist Fatalities", "Total Motorcyclists")) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+```
+\newpage
+
+```{r}
+#| label: fig-hour-motorcycle
+#| fig-cap: Number of Fatalities by Hour for Motorcyclists
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-width: 5
+#| fig-height: 2.5
+#| fig-align: center
+
+# Create interaction plot data for motorcycle
+interaction_motorcycle <- analysis_data %>%
+  filter(motorcycle == 1) %>%
+  group_by(hour) %>%
+  summarize(motorcycle_fatalities = sum(fatalities == 1, na.rm = TRUE))
+
+# Plot the interaction between hour and motorcycle
+ggplot(interaction_motorcycle, aes(x = hour, y = motorcycle_fatalities)) +
+  geom_line(color = "red", size = 1.2) +
+  labs(x = "Hour of Day", y = "Motorcycle Fatalities") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
 ```
@@
 #| label: fig-hour-pedestrian
-#| fig-cap: Fatality Likelihood by Hour for Pedestrians
+#| fig-cap: Number of Fatalities by Hour for Pedestrians
 #| echo: false
@@
 summary_stats_pedestrian %>%
-  kable() %>%
+  kable(col.names = c("Hour", "Pedestrian Fatalities", "Total Pedestrians")) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
@@
 
-\begin{align}S_i \sim \text{Bernoulli}(p_i)\end{align}
-
-\newpage
+\begin{align}Y_i \sim \text{Bernoulli}(p_i)\end{align}
 
@@
 
-The model was implemented in R [@citeR] using the `rstanarm` package, which offers an accessible interface for Bayesian generalized linear models (GLMs), allowing specification of priors and customization of model parameters.
+The model was implemented in R [@citeR] using the `rstanarm` package, which provides an accessible interface for Bayesian generalized linear models (GLMs), allowing specification of priors and customization of model parameters.
 
@@
 The model assumes independence of observations, meaning that each observation's outcome is not influenced by others.
-However, this assumption may be partially violated in the context of vehicle accidents, as incidents occurring within close temporal or spatial proximity may share common influencing factors, such as weather, road conditions, or traffic patterns.
+However, this assumption may be partially violated in the context of vehicle accidents, as incidents occurring within close temporal or spatial proximity may share common influencing variables, such as weather, road conditions, or traffic patterns.
 While the inclusion of temporal and spatial variables like hour and police division aims to mitigate such dependencies, it may not fully account for potential clustering effects.
@@
 
-Another important limitation is the potential for omitted variable bias and unmeasured confounding factors.
+Another important limitation is the potential for omitted variable bias and unmeasured confounding variables.
 For example, the model does not include variables like road infrastructure quality, driver impairment (e.g., alcohol or drug use), or real-time weather conditions, all of which could influence accident outcomes.
@@
 Additionally, the effective sample size $n_{\text{eff}}$ was high across all parameters, suggesting low autocorrelation and contributing to model stability.
-See more details of our model diagnostics here: (INSERT REFERENCE)
-
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| tbl-align: center
-#| fig-height: 10
-#| label: tbl-modelresults
-#| tbl-cap: "Summary for Motor Vehicle Collision Fatality Prediction Model"
-
-modelsummary(
-  prediction_model,
-  output = "kableExtra",
-) 
-```
-@tbl-modelresults presents the estimated coefficients for the predictors in our GLM model.
+See more details of our model diagnostics in @sec-appendix-diagnostics.
+
+@tbl-modelresults1 and @tbl-modelresults2 presents the estimated coefficients for the predictors in our GLM model.
 These coefficients fit into the GLM equation, \ref{eq:logit}, allowing us to interpret the impact of each predictor on the likelihood of fatality in a vehicular accident.
@@
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-#| fig-align: center
-#| label: fig-coefficient-estimates
-#| fig-height: 10
-#| fig-cap: "Coefficient Estimates for Predictors"
-
-# Tidy Model Summary Data
-coefficients <- broom.mixed::tidy(prediction_model, conf.int = TRUE) %>%
-  mutate(across(c(estimate, std.error, conf.low, conf.high), round, 2))
-
-# Make Plot
-coefficients %>%
-  ggplot(aes(estimate, term)) + 
-  geom_point() + 
-  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) + 
-  labs(x = "Estimate", y = "Predictor") +
-  theme(axis.text.y = element_text(size = 7))
-```
-
 @fig-coefficient-estimates represents the model coefficients, with error bars indicating the confidence interval for each estimate.
 While most variables are relatively in line, it seems that the interaction terms all have wide confidence intervals, suggesting high variability in its estimates.
+All plots can be found in @sec-appendix-results.
 
@@
 Specifically, the high risk faced by motorcyclists, pedestrians, and individuals involved in accidents during rush hours or late evenings aligns with long-held perceptions about these vulnerable groups.
-However, the study also introduces a nuanced understanding of these factors, suggesting deeper complexities that could inform public policy.
+However, the study also introduces a nuanced understanding of these variables, suggesting deeper complexities that could inform public policy.
 
@@
 
-In @tbl-modelresults, the coefficient of 17.036 for motorcyclists indicates that they are the most at-risk group in terms of MVC fatalities.
+In @tbl-modelresults1 and @tbl-modelresults2, the coefficient of 17.036 for motorcyclists indicates that they are the most at-risk group in terms of MVC fatalities.
 This result strongly aligns with the widely recognized vulnerability of motorcyclists in traffic accidents.
-Motorcycles, by their very design, offer little protection compared to other vehicles.
+Motorcycles, by their very design, providing little protection compared to other vehicles.
 In the event of a crash, motorcyclists are more likely to suffer severe injuries due to the lack of structural safeguards such as airbags, seat belts, or the protective frame that cars provide.
@@
 
-Similarly, shown in @tbl-modelresults, the coefficient of 6.809 for pedestrians reinforces the widely acknowledged dangers pedestrians face in urban settings.
-Pedestrian safety has long been a critical area of concern for urban planners and policymakers, and this result highlights the need for continued focus on this issue.
+Similarly, shown in @tbl-modelresults1, the coefficient of 6.809 for pedestrians reinforces the widely acknowledged dangers pedestrians face in urban settings.
+Pedestrian safety has long been a key area of concern for urban planners and policymakers, and this result highlights the need for continued focus on this issue.
 Pedestrians are inherently vulnerable in any accident, and the lack of physical barriers between them and vehicles significantly increases their chances of injury or death in a collision.
@@
 
-### Environmental Factors
-
-The analysis also highlights the significant role that environmental factors—specifically, the time of day—play in MVC fatalities.
+### Environmental Variables
+
+The analysis also highlights the significant role that environmental variables—specifically, the time of day—play in MVC fatalities.
 During rush hours (particularly between hours 16 and 17) and late evenings (around 11:00 PM), the likelihood of a fatality, especially in motorcycle crashes, increases substantially.
-The 66% fatality likelihood for motorcycle crashes during these times suggests that several environmental and behavioral factors may be at play.
+The 66% fatality likelihood for motorcycle crashes during these times suggests that several environmental and behavioral elements may be at play.
 During rush hours, the density of traffic can lead to higher stress levels among drivers, potentially contributing to reckless driving, lane changes, or a lack of attention to motorcyclists.
@@
 
-## Role of Protective Factors {#sec-discussion-protective-factors}
-
-The variables "injury_collision" and "property damage collisions" play a crucial role in lowering the likelihood of fatalities in motor vehicle collisions (MVCs).
+## Role of Type of Crash Variables {#sec-discussion-crash-variables}
+
+The variables "injury_collision" and "property damage collisions" play an important role in lowering the likelihood of fatalities in motor vehicle collisions (MVCs).
 These outcomes suggest that, in contrast to fatal accidents, certain collisions are characterized by lower severity, different dynamics, and an overall reduced risk to life.
-By exploring why these variables lower fatality risks, we can better understand the factors that contribute to non-fatal collisions and how interventions might be designed to replicate these protective conditions across all collisions.
+By exploring why these variables lower fatality risks, we can better understand the elements that contribute to non-fatal collisions and how interventions might be designed to replicate these protective conditions across all collisions.
 
@@
 This quick medical intervention can prevent fatalities by addressing life-threatening injuries before they become fatal.
-Analyzing injury collisions can therefore offer valuable insights into what makes these accidents less deadly and how certain safety features, such as improved emergency response times, advanced vehicle safety systems, and better road infrastructure, can prevent fatalities.
-
-Property damage collisions are another protective factor that lowers the likelihood of fatalities.
+Analyzing injury collisions can therefore provide perspectives into what makes these accidents less deadly and how certain safety features, such as improved emergency response times, advanced vehicle safety systems, and better road infrastructure, can prevent fatalities.
+
+Property damage collisions are another variable that lowers the likelihood of fatalities.
 These types of collisions, by definition, involve no physical harm to the individuals involved, but they still represent significant risks of damage to the vehicles and the surrounding environment.
@@
 
-While the results of the analysis provide valuable insights into the factors influencing MVC fatalities, they also highlight significant uncertainties, particularly with some predictors such as the interaction terms.
+While the results of the analysis provide perspectives into the elements influencing MVC fatalities, they also highlight significant uncertainties, particularly with some predictors such as the interaction terms.
 In @fig-coefficient-estimates, the variability of the interaction terms underscores the inherent challenges of modeling rare events such as fatalities, where small changes in conditions can lead to disproportionately large impacts.
@@
 
-The variability in the coefficients for predictors l suggests that there may be additional, unobserved factors influencing the likelihood of fatalities that are not captured by the current model.
+The variability in the coefficients for predictors l suggests that there may be additional, unobserved variables influencing the likelihood of fatalities that are not captured by the current model.
 This uncertainty could stem from gaps in the available data or from the limitations in the types of variables collected, such as the lack of detailed road conditions, driver behavior, or vehicle-specific characteristics (e.g., model, age of the vehicle).
 To reduce this uncertainty, a more comprehensive data collection strategy is needed, focusing on variables that could provide a deeper understanding of accident dynamics, such as weather conditions, road infrastructure quality, or even the socio-economic status of the areas where accidents occur.
-By expanding the scope and granularity of the data, it would be possible to refine the model, improving the precision of the estimates and providing clearer insights into the factors that contribute to fatalities.
-
-The variability in the model’s predictions also speaks to the stochastic nature of accidents, where the outcome of any given crash is influenced by a multitude of unpredictable factors.
+By expanding the scope and granularity of the data, it would be possible to refine the model, improving the precision of the estimates and providing clearer perspectives into the variables that contribute to fatalities.
+
+The variability in the model’s predictions also speaks to the stochastic nature of accidents, where the outcome of any given crash is influenced by a multitude of unpredictable elements.
 MVC fatalities, while significant, are relatively rare compared to non-fatal accidents, making them inherently difficult to predict.
 In this context, the uncertainty observed in the model's results can be understood as a reflection of the randomness and unpredictability involved in traffic collisions.
-Factors such as driver response times, unexpected road hazards, or even the randomness of the sequence of events leading up to a fatal crash are beyond the scope of even the most sophisticated predictive models.
+Variables such as driver response times, unexpected road hazards, or even the randomness of the sequence of events leading up to a fatal crash are beyond the scope of even the most sophisticated predictive models.
 In addition, fatalities often result from a confluence of circumstances—such as high speed, poor weather conditions, or the failure of safety mechanisms—that may not be fully captured by existing data or model specifications.
@@
 
-The findings from this analysis provide valuable insights for policymakers and law enforcement agencies seeking to improve road safety.
+The findings from this analysis provide lessons for policymakers and law enforcement agencies seeking to improve road safety.
 One of the most pressing implications is the identification of high-risk periods, particularly during rush hours and late evenings, where the likelihood of fatal collisions is significantly higher.
-Targeted interventions could focus on enforcing speed limits and impaired driving laws during these critical hours, which may help mitigate the increased risk of accidents.
+Targeted interventions could focus on enforcing speed limits and impaired driving laws during these key hours, which may help mitigate the increased risk of accidents.
 For example, police could increase patrols and sobriety checkpoints, particularly in areas with higher concentrations of motorcycle or pedestrian collisions.
@@
 These measures could help reduce the exposure of vulnerable road users to traffic hazards, particularly during high-risk times.
-Furthermore, predictive insights from statistical modeling could revolutionize real-time road safety strategies.
+Furthermore, statistical modeling could revolutionize real-time road safety strategies.
 For instance, traffic management systems could use predictive data to adjust traffic light patterns during peak hours, optimizing traffic flow and reducing congestion-related accidents.
-Additionally, emergency services could use data-driven insights to allocate resources more efficiently, ensuring that areas with higher likelihoods of fatal collisions are prioritized for immediate response.
+Additionally, emergency services could use quantitative data to allocate resources more efficiently, ensuring that areas with higher likelihoods of fatal collisions are prioritized for immediate response.
 In this way, statistical modeling can provide law enforcement and urban planners with the tools necessary to not only react to but proactively address high-risk situations on the road.
@@
 The model’s limitations stem from both the stochastic nature of traffic accidents and potential biases in the data.
-Fatalities are rare events, and the randomness inherent in accident outcomes—such as variations in road conditions, driver behavior, or other small factors—makes it difficult for models to predict them with high precision.
-Additionally, biases in data collection, like inconsistent reporting of certain variables or underreporting of less severe accidents, may lead to seleciton bias towards more severe accidnets in our datasets, affecting our predictions.
-For example, in the United States, almost a third of car crashes involving a vulnerable road user go unreported (INSERT CITATION).
+Fatalities are rare events, and the randomness inherent in accident outcomes—such as variations in road conditions, driver behavior, or other small variables—makes it difficult for models to predict them with high precision.
+Additionally, biases in data collection, like inconsistent reporting of certain variables or underreporting of less severe accidents, may lead to seleciton bias towards more severe accidents in our datasets, affecting our predictions.
+For example, in the United States, almost a third of car crashes involving a vulnerable road user go unreported [@streetsblog2021].
 Though this is a statistics from the United States, the article's findings highlight that the potential biases present in our data.
@@
 Moreover, the model’s reliance on aggregate data limits its ability to account for the nuanced, dynamic nature of real-time road events.
-The factors influencing a fatality may not be static and can vary significantly based on the specific context of each collision.
-This variability makes it challenging to create a one-size-fits-all model for MVC fatalities, and the results are likely to change with shifts in environmental, societal, or technological factors.
+The elements influencing a fatality may not be static and can vary significantly based on the specific context of each collision.
+This variability makes it challenging to create a one-size-fits-all model for MVC fatalities, and the results are likely to change with shifts in environmental, societal, or technological variables.
 To improve accuracy and robustness, further data collection is essential, particularly to capture rare but impactful events and refine predictors of fatality outcomes.
-Enhancements in data quality and the incorporation of more granular, real-time data will improve model predictions and provide better insight into the complex dynamics of traffic accidents.
-
-As a next step, future research should focus on collecting more granular, real-time data to capture the dynamic factors influencing MVC fatalities.
+Enhancements in data quality and the incorporation of more granular, real-time data will improve model predictions and help understand the complex dynamics of traffic accidents.
+
+As a next step, future research should focus on collecting more granular, real-time data to capture the dynamic elements influencing MVC fatalities.
 This could involve leveraging emerging technologies such as smart traffic sensors, vehicle telematics, and data from wearable devices to capture more detailed and context-specific information.
 Additionally, refining statistical techniques, such as Bayesian hierarchical models or machine learning approaches, can help better account for the inherent randomness of accidents and the uncertainty in predicting rare outcomes.
-Improved modeling could lead to more accurate, actionable insights for policy development and real-time road safety interventions.
+Improved modeling could lead to more accurate, actionable perspectives for policy development and real-time road safety interventions.
 
@@
 
-## Exploration of Surveys, Sampling, and Observational Data {#sec-appendix-sso}
+## Surveys, Sampling, and Observational Data {#sec-appendix-sso}
 
@@
 The Toronto Police Service's Motor Vehicle Collision database is collected using a systematic process.
-When a collision occurs such that where persons are injured or a combined damage is valued at more than $\2000 to vehicles or property, or damage to any private, municipal or highway property, police officers complete a standardized Motor Vehicle Accident Report form.
+When a collision occurs such that where persons are injured or a combined damage is valued at more than \$2000 to vehicles or property, or damage to any private, municipal or highway property, police officers complete a standardized Motor Vehicle Accident Report form.
 This data is then maintained by the City of Toronto, where the collection process follow a standardized process highlighted in the Motor Vehicle Collision Report manual [@mvcr_guide], ensuring reporting consistency.
@@
 
-The current data collection methodology employed by the Toronto Police Service provides valuable insights into Motor Vehicle Collisions (MVCs), yet several enhancements could make the dataset more comprehensive and actionable.
-Expanding the criteria for reporting is a critical first step.
+The current data collection methodology employed by the Toronto Police Service provides perspectives into Motor Vehicle Collisions (MVCs), yet several enhancements could make the dataset more comprehensive and actionable.
+Expanding the criteria for reporting is an important first step.
 By including minor collisions that do not meet the $2,000 damage threshold or require police presence, the city can gain a fuller understanding of road safety dynamics.
@@
 Integration of traffic camera data, analyzed using AI-powered algorithms, would add an unbiased and continuous data stream to validate collision reports and identify patterns.
-These methods collectively address gaps in the current system, providing deeper insights into the factors contributing to MVCs.
+These methods collectively address gaps in the current system, providing deeper lessons into the elements contributing to MVCs.
 
 Expanding the scope of variables collected in collision reports is another avenue for improvement.
-Adding details about driver behavior (e.g., speeding or distracted driving), the involvement of emerging vehicle technologies (e.g., self-driving systems), and the road environment (e.g., construction zones) can offer richer insights into the conditions that contribute to MVCs.
-This would help policymakers and researchers better understand the interplay of human, environmental, and technological factors in collision outcomes.
+Adding details about driver behavior (e.g., speeding or distracted driving), the involvement of emerging vehicle technologies (e.g., self-driving systems), and the road environment (e.g., construction zones) can provide new perspectives into the conditions that contribute to MVCs.
+This would help policymakers and researchers better understand the interplay of human, environmental, and technological variables in collision outcomes.
 
@@
 
-To gather more information about motor vehicle collisions, I've attached a proposed survey to better understand the factors that contribute to fatalities and injuries.
+To gather more information about motor vehicle collisions, I've attached a proposed survey to better understand the items that contribute to fatalities and injuries.
 We aim to replicate something similar to the Motor Vehicle Collision Report and aim to collect data for car crashes in Toronto.
@@
 ### Model Overview {#sec-appendix-modelcard-overview}
-This model predicts the occurrence of motor vehicle collisions (MVCs) based on various factors, including the time of day (hour), type of vehicle involved (motorcycle, pedestrian, automobile, etc.), and police division.
+This model predicts the occurrence of motor vehicle collisions (MVCs) based on various elements, including the time of day (hour), type of vehicle involved (motorcycle, pedestrian, automobile, etc.), and police division.
 The model is trained using data from 2015 to 2024, with a focus on the interactions between the hour of the day and motorcycle or pedestrian involvement.
@@
 ### Conclusion {#sec-appendix-modelcard-conclusion}
-The model performs exceptionally well in predicting motor vehicle collisions based on hour, vehicle type, and other factors, with the highest predictive power for motorcycle-related accidents.
+The model performs exceptionally well in predicting motor vehicle collisions based on hour, vehicle type, and other variables, with the highest predictive power for motorcycle-related accidents.
 Key interactions, such as between hour and motorcycle or pedestrian, highlight the importance of specific hours of the day in predicting collision occurrences.
@@
 The close alignment suggests that the model accurately captures the data's central tendency and variability.
-@fig-convergencecheckTrace and @fig-convergencecheckRhat show that the sampling algorithm used, the Markov chain Monte Carlo (MCMC) algorithm, did not run into issues as the posterior distribution for the model was created.
+@fig-convergencecheckTrace, and @fig-convergencecheckRhat show that the sampling algorithm used, the Markov chain Monte Carlo (MCMC) algorithm, did not run into issues as the posterior distribution for the model was created.
 Using the checks presented by @tellingstories, both graphs do not show anything abnormal since the trace plots in @fig-convergencecheckTrace display substantial horizontal fluctuation across chains, indicating good mixing, while the Rhat values in @fig-convergencecheckRhat are close to 1 and well below 1.05, further supporting convergence.
@@
 #| warning: false
+#| fig-height: 15
+#| fig-width: 20
 #| label: fig-convergencecheckTrace
 #| fig-cap: "Checking the convergence of the MCMC algorithm - Trace"
-trace_plot <- plot(prediction_model, "trace") + theme_minimal()
-grid.arrange(trace_plot, ncol = 5, nrow = 5)
+
+# Generate the trace plot
+plot(prediction_model, "trace") + theme_minimal()
 ```
@@
 #| warning: false
+#| fig-height: 13
 #| label: fig-convergencecheckRhat
@@
 ```
+## Model Results {#sec-appendix-results}
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| tbl-align: center
+#| fig-height: 10
+#| label: tbl-modelresults1
+#| tbl-cap: "Summary for Motor Vehicle Collision Fatality Prediction Model Part 1"
+
+coefficients <- broom.mixed::tidy(prediction_model, conf.int = TRUE) %>%
+  mutate(across(c(estimate, std.error, conf.low, conf.high), round, 2))
+
+# Calculate the number of rows and midpoint
+midpoint <- ceiling(nrow(coefficients) / 2)
+
+# Split the dataframe into two parts
+coefficients_part1 <- coefficients[1:midpoint, ]
+coefficients_part2 <- coefficients[(midpoint + 1):nrow(coefficients), ]
+
+# Generate the first table
+kable(coefficients_part1, format = "latex", booktabs = TRUE, align = "c") %>% 
+  kable_styling(latex_options = "scale_down", font_size = 7)
+
+```
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| tbl-align: center
+#| fig-height: 10
+#| label: tbl-modelresults2
+#| tbl-cap: "Summary for Motor Vehicle Collision Fatality Prediction Model Part 2"
+
+coefficients <- broom.mixed::tidy(prediction_model, conf.int = TRUE) %>%
+  mutate(across(c(estimate, std.error, conf.low, conf.high), round, 2))
+
+# Calculate the number of rows and midpoint
+midpoint <- ceiling(nrow(coefficients) / 2)
+
+# Split the dataframe into two parts
+coefficients_part1 <- coefficients[1:midpoint, ]
+coefficients_part2 <- coefficients[(midpoint + 1):nrow(coefficients), ]
+
+# Generate the second table
+kable(coefficients_part2, format = "latex", booktabs = TRUE, align = "c") %>% 
+  kable_styling(latex_options = "scale_down", font_size = 6)
+
+```
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-align: center
+#| label: fig-coefficient-estimates
+#| fig-height: 10
+#| fig-cap: "Coefficient Estimates for Predictors"
+
+# Tidy Model Summary Data
+coefficients <- broom.mixed::tidy(prediction_model, conf.int = TRUE) %>%
+  mutate(across(c(estimate, std.error, conf.low, conf.high), round, 2))
+
+# Make Plot
+coefficients %>%
+  ggplot(aes(estimate, term)) + 
+  geom_point() + 
+  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) + 
+  labs(x = "Estimate", y = "Predictor") +
+  theme(axis.text.y = element_text(size = 7))
+```
 

### 2024-12-04 paper/paper.qmd
adjusted trace plots

@@
 #| warning: false
-#| fig-height: 15
-#| fig-width: 20
+#| fig-height: 13
+#| fig-width: 10
 #| label: fig-convergencecheckTrace
@@
 
-# Generate the trace plot
-plot(prediction_model, "trace") + theme_minimal()
+# Had to create my own trace plots because I couldn't adjust the formula
+# Extract parameter names
+trace_data <- as.data.frame(as.matrix(prediction_model))
+params <- colnames(trace_data)
+param_chunks <- split(params, ceiling(seq_along(params) / 20))
+
+# Function to plot trace plots
+plot_trace <- function(params_subset, data, title) {
+  data %>%
+    select(all_of(params_subset)) %>%
+    pivot_longer(cols = everything(), names_to = "Parameter", values_to = "Value") %>%
+    ggplot(aes(x = rep(1:nrow(data), length(params_subset)), y = Value, color = Parameter)) +
+    geom_line() +
+    facet_wrap(~ Parameter, scales = "free", ncol = 4) +  
+    theme_minimal() +
+    labs(title = title, x = "Iteration", y = "Value") +
+    theme(legend.position = "none") + 
+    theme(
+      plot.margin = margin(t = 5, r = 5, b = 5, l = 10),  
+      axis.text.x = element_text(size = 5, angle = 45, hjust = 1), 
+      axis.text.y = element_text(size = 5), 
+      plot.title = element_text(size = 10, hjust = 0.5), 
+      strip.text = element_text(size = 7))
+}
+
+# Generate individual trace plots for each chunk
+trace_plots <- list()
+for (i in seq_along(param_chunks)) {
+  trace_plots[[i]] <- plot_trace(
+    params_subset = param_chunks[[i]],
+    data = trace_data,
+    title = paste("Trace Plots (Part", i, ")")
+  )
+}
+
+grid.arrange(grobs = trace_plots, ncol = 2)
 ```
@@
 #| fig-height: 13
+#| fig-width: 10
 #| label: fig-convergencecheckRhat
@@
 plot(prediction_model, "rhat") + theme_minimal()
-```
+
+```
+\newpage
+
 ## Model Results {#sec-appendix-results}
+
+Please find the results to our model in @tbl-modelresults1, @tbl-modelresults2, and @fig-coefficient-estimates.
 

### 2024-12-04 paper/paper.qmd
final edits

@@
 By analyzing the interaction between hour of the day and vehicle types, including motorcycles and pedestrians, the model identifies key elements contributing to fatal outcomes.
-Results indicate that late-afternoon to evening hours, particularly between 4 PM and 11 PM, show a higher likelihood of fatalities, with motorcycles and pedestrians significantly increasing fatality risk.
+Results indicate that late-afternoon to evening hours, particularly between 5 PM and 11 PM, show a higher likelihood of fatalities, with motorcycles and pedestrians significantly increasing fatality risk.
 The model achieved a high predictive accuracy, demonstrating its potential in informing safety policies and interventions in Toronto."
@@
 
-Our analysis identifies key predictors of vehicular accident fatalities, emphasizing heightened risks for motorcyclists and pedestrians, especially during rush hours (4PM to 5PM) and late evenings (11:00 PM).
-Motorcycle collisions in the evening show a fatality likelihood of around 66%.
+Our analysis identifies key predictors of vehicular accident fatalities, emphasizing heightened risks for motorcyclists and pedestrians, especially during rush hour and late evenings (11:00 PM).
+Motorcycle collisions in the evening show a fatality likelihood of around 53%.
 Collision type variables, such as injury and property damage collisions, significantly lower fatality risks, while police division and year have minimal impact.
@@
 These coefficients fit into the GLM equation, \ref{eq:logit}, allowing us to interpret the impact of each predictor on the likelihood of fatality in a vehicular accident.
-Most notably, motorcyclist (17.036) and pedestrian (6.809) collisions exhibit substantial positive associations with fatality occurrence, which is in line with expectations.
-Moreover, motorcyclist accidents during rush hour (hour 16 and 17) have the highest coefficients at 40.4 and 26.8, respectively, and during the late evening, such as at 11:00 PM (39.365), indicating that a collision involving a motorcycle during rush hour or late evening is around 66\% likely to result in a fatality.
+Most notably, motorcyclist (10.319) and pedestrian (6.766) collisions exhibit substantial positive associations with fatality occurrence, which is in line with expectations.
+Moreover, motorcyclist accidents during rush hour (5 PM) have the highest coefficients at 23.656, and during the late evening, such as at 11:00 PM (37.984), indicating that a collision involving a motorcycle during rush hour or late evening is around 53\% likely to result in a fatality.
 We also find similar results with pedestrians as well.
@@
 
-In @tbl-modelresults1 and @tbl-modelresults2, the coefficient of 17.036 for motorcyclists indicates that they are the most at-risk group in terms of MVC fatalities.
+In @tbl-modelresults1 and @tbl-modelresults2, the coefficient of 10.319 for motorcyclists indicates that they are the most at-risk group in terms of MVC fatalities.
 This result strongly aligns with the widely recognized vulnerability of motorcyclists in traffic accidents.
@@
 The analysis also highlights the significant role that environmental variables—specifically, the time of day—play in MVC fatalities.
-During rush hours (particularly between hours 16 and 17) and late evenings (around 11:00 PM), the likelihood of a fatality, especially in motorcycle crashes, increases substantially.
-The 66% fatality likelihood for motorcycle crashes during these times suggests that several environmental and behavioral elements may be at play.
+During rush hour and late evenings (around 11:00 PM), the likelihood of a fatality, especially in motorcycle crashes, increases substantially.
+The 53% fatality likelihood for motorcycle crashes during these times suggests that several environmental and behavioral elements may be at play.
 During rush hours, the density of traffic can lead to higher stress levels among drivers, potentially contributing to reckless driving, lane changes, or a lack of attention to motorcyclists.
@@
 
-- **Intercept**: 9.705
+- **Intercept**: 13.146
 - **Hour Coefficients**:
@@
   - **Key Hours**:
-    - `hour2`: 1.920
-    - `hour3`: 2.908
-    - `hour6`: 0.966
-    - `hour7`: -1.673
-    - `hour8`: -0.511
-    - `hour21`: 1.237
-    - `hour22`: 0.631
-    - `hour23`: 1.572
+    - `hour2`: 1.449
+    - `hour3`: 3.851
+    - `hour6`: 0.350
+    - `hour7`: 0.233
+    - `hour8`: 0.080
+    - `hour21`: -0.184
+    - `hour22`: 0.821
+    - `hour23`: 2.342
 - **Vehicle Type Coefficients**:
-  - **Motorcycle**: 17.036, indicating a high likelihood of motorcycle-related collisions.
-  - **Pedestrian**: 6.832, showing the significance of pedestrian involvement in accidents.
-  - **Bicycle**: 3.809, showing a moderate likelihood of bicycle-related accidents.
-  - **Automobile**: 2.648, indicating that automobile-related collisions are also common.
+  - **Motorcycle**: 10.32, indicating a high likelihood of motorcycle-related collisions.
+  - **Pedestrian**: 6.77, showing the significance of pedestrian involvement in accidents.
+  - **Bicycle**: 3.45, showing a moderate likelihood of bicycle-related accidents.
+  - **Automobile**: -0.26 , indicating that automobile-related collisions are also common.
 - **Collision Type Coefficients**:
@@
 - **Police Division Coefficients**:
-  - `police_divisionD12`: 0.846, indicating division-specific effects, where division D12 has a positive association with collisions.
-  - `police_divisionD13`: 0.075
-  - `police_divisionD14`: -2.806
-  - `police_divisionD22`: 1.797
-  - `police_divisionD23`: 1.193
-  - `police_divisionD31`: 0.619
-  - `police_divisionD32`: 1.007
-  - `police_divisionD33`: 0.055
-  - `police_divisionD41`: 0.989
-  - `police_divisionD42`: 1.140
-  - `police_divisionD43`: 0.582
-  - `police_divisionD51`: 0.915
-  - `police_divisionD52`: 1.465
-  - `police_divisionD53`: 1.389
-  - `police_divisionD55`: 0.389
-  - `police_divisionNSA`: -4.040
+  - `police_divisionD12`: 0.06, indicating division-specific effects, where division D12 has a positive association with collisions.
+  - `police_divisionD13`: 0.17
+  - `police_divisionD14`: -1.00
+  - `police_divisionD22`: -0.09
+  - `police_divisionD23`: 0.93
+  - `police_divisionD31`: -0.65
+  - `police_divisionD32`: 0.21
+  - `police_divisionD33`: 0.14
+  - `police_divisionD41`: 0.12
+  - `police_divisionD42`: 0.46
+  - `police_divisionD43`: 0.38
+  - `police_divisionD51`: -0.34
+  - `police_divisionD52`: 4.78
+  - `police_divisionD53`: -0.88
+  - `police_divisionD55`: -0.93
+  - `police_divisionNSA`: -3.24
 - **Year Coefficients**:
-  - `year2015`: -0.500
-  - `year2016`: 0.329
-  - `year2017`: 0.412
-  - `year2018`: 0.165
-  - `year2019`: -0.025
-  - `year2020`: 0.021
-  - `year2021`: 0.079
-  - `year2022`: 1.947, showing a significant increase in collisions during this year.
-  - `year2023`: -1.708, showing a decrease in collisions during this year.
-  - `year2024`: -0.210
+  - `year2015`: 1.03
+  - `year2016`: 1.47
+  - `year2017`: 1.18
+  - `year2018`: 0.93
+  - `year2019`: 0.55
+  - `year2020`: 0.61
+  - `year2021`: 1.85
+  - `year2022`: 3.17, showing a significant increase in collisions during this year.
+  - `year2023`: 0.14
+  - `year2024`: 0.75
 - **Interaction Terms**:
   - **Hour × Motorcycle**:
-    - `hour1 × motorcycle`: 31.900
-    - `hour2 × motorcycle`: 33.094
-    - `hour3 × motorcycle`: 33.377
-    - `hour4 × motorcycle`: 33.633
-    - `hour6 × motorcycle`: 51.053
-    - `hour7 × motorcycle`: 52.055
-    - `hour8 × motorcycle`: 55.746
-    - `hour9 × motorcycle`: -66.051
-    - `hour10 × motorcycle`: 27.802
-    - `hour12 × motorcycle`: 51.049
-    - `hour13 × motorcycle`: 18.620
-    - `hour14 × motorcycle`: -12.621
-    - `hour15 × motorcycle`: -11.489
-    - `hour16 × motorcycle`: 32.651
-    - `hour17 × motorcycle`: 20.150
-    - `hour18 × motorcycle`: 12.044
-    - `hour19 × motorcycle`: -11.899
-    - `hour20 × motorcycle`: -5.807
-    - `hour21 × motorcycle`: 21.915
-    - `hour22 × motorcycle`: 9.369
-    - `hour23 × motorcycle`: 34.897
+    - `hour1 × motorcycle`: 33.237
+    - `hour2 × motorcycle`: 35.448
+    - `hour3 × motorcycle`: 35.467
+    - `hour4 × motorcycle`: 36.943
+    - `hour6 × motorcycle`: 57.233
+    - `hour7 × motorcycle`: -6.889
+    - `hour8 × motorcycle`: 9.962
+    - `hour10 × motorcycle`: 29.521
+    - `hour12 × motorcycle`: 54.822
+    - `hour13 × motorcycle`: 19.375
+    - `hour14 × motorcycle`: 24.927
+    - `hour15 × motorcycle`: 23.135
+    - `hour16 × motorcycle`: -5.637
+    - `hour17 × motorcycle`: 23.656
+    - `hour18 × motorcycle`: -5.293
+    - `hour19 × motorcycle`: 24.156
+    - `hour20 × motorcycle`: -2.714
+    - `hour21 × motorcycle`: -3.445
+    - `hour22 × motorcycle`: 11.598
+    - `hour23 × motorcycle`: 37.984
   - **Hour × Pedestrian**:
-    - `hour1 × pedestrian`: 15.095
-    - `hour2 × pedestrian`: 29.594
-    - `hour3 × pedestrian`: 20.771
-    - `hour4 × pedestrian`: 19.986
-    - `hour5 × pedestrian`: 23.943
-    - `hour6 × pedestrian`: 9.778
-    - `hour7 × pedestrian`: 1.434
-    - `hour8 × pedestrian`: 0.484
-    - `hour9 × pedestrian`: 1.831
-    - `hour10 × pedestrian`: 1.895
-    - `hour11 × pedestrian`: 15.280
-    - `hour12 × pedestrian`: 9.497
-    - `hour13 × pedestrian`: 1.397
-    - `hour14 × pedestrian`: 12.171
-    - `hour15 × pedestrian`: -0.552
-    - `hour16 × pedestrian`: 0.453
-    - `hour17 × pedestrian`: 0.272
-    - `hour18 × pedestrian`: -5.210
-    - `hour19 × pedestrian`: -0.707
-    - `hour20 × pedestrian`: 7.720
-    - `hour21 × pedestrian`: 12.510
-    - `hour22 × pedestrian`: 1.760
-    - `hour23 × pedestrian`: -2.690
+    - `hour1 × pedestrian`: 15.758
+    - `hour2 × pedestrian`: 29.368
+    - `hour3 × pedestrian`: 18.546
+    - `hour4 × pedestrian`: 20.723
+    - `hour5 × pedestrian`: 22.835
+    - `hour6 × pedestrian`: 1.713
+    - `hour7 × pedestrian`: 17.229
+    - `hour8 × pedestrian`: -2.876
+    - `hour9 × pedestrian`: 0.983
+    - `hour10 × pedestrian`: 0.973
+    - `hour11 × pedestrian`: 9.320
+    - `hour12 × pedestrian`: -2.071
+    - `hour13 × pedestrian`: -1.027
+    - `hour14 × pedestrian`: 12.206
+    - `hour15 × pedestrian`: -0.145
+    - `hour16 × pedestrian`: 0.042
+    - `hour17 × pedestrian`: -0.494
+    - `hour18 × pedestrian`: -0.977
+    - `hour19 × pedestrian`: 0.090
+    - `hour20 × pedestrian`: 1.757
+    - `hour21 × pedestrian`: -0.436
+    - `hour22 × pedestrian`: 14.945
+    - `hour23 × pedestrian`: 12.795
 
@@
 - **R²**: 0.992, indicating excellent fit to the data.
-- **Log Likelihood**: -0.082
-- **Effective Log Predictive Density (ELPD)**: -16.2, with an associated standard error of 2.2.
-- **Leave-One-Out Information Criterion (LOOIC)**: 32.4, with a standard error of 4.4.
+- **Log Likelihood**: -0.326
+- **Effective Log Predictive Density (ELPD)**: -17.6, with an associated standard error of 3.1.
+- **Leave-One-Out Information Criterion (LOOIC)**: 35.1, with a standard error of 6.2.
 - **Watanabe-Akaike Information Criterion (WAIC)**: 21.3
@@
 - **Chain 1**:
-  - Warm-up: 29.99 seconds
-  - Sampling: 26.11 seconds
-  - Total: 56.1 seconds
+  - Warm-up: 24.63 seconds
+  - Sampling: 18.63 seconds
+  - Total: 43.26 seconds
 - **Chain 2**:
-  - Warm-up: 19.62 seconds
-  - Sampling: 17.23 seconds
-  - Total: 36.84 seconds
+  - Warm-up: 22.78 seconds
+  - Sampling: 26.92 seconds
+  - Total: 49.70 seconds
 - **Chain 3**:
-  - Warm-up: 19.88 seconds
-  - Sampling: 19.30 seconds
-  - Total: 39.18 seconds
+  - Warm-up: 31.42 seconds
+  - Sampling: 24.05 seconds
+  - Total: 55.47 seconds
 - **Chain 4**:
-  - Warm-up: 22.40 seconds
-  - Sampling: 19.17 seconds
-  - Total: 41.57 seconds
+  - Warm-up: 26.75 seconds
+  - Sampling: 27.47 seconds
+  - Total: 54.21 seconds
 

### 2024-12-04 paper/paper.qmd
updated model card

@@
 - **Hour and Motorcycle**:
-  - `hour1 × motorcycle`: 31.900, significant interaction in early morning hours, suggesting motorcycle accidents are more likely during certain hours.
-  - Other significant hours for motorcycle collisions are `hour6 × motorcycle` (51.053) and `hour9 × motorcycle` (-66.051).
+  - `hour1 × motorcycle`: 33.237, significant interaction in early morning hours, suggesting motorcycle accidents are more likely during certain hours.
+  - Other significant hours for motorcycle collisions are `hour12 × motorcycle` (54.822) and `hour23 × motorcycle` (37.984).
 - **Hour and Pedestrian**:
-  - `hour1 × pedestrian`: 15.095, significant pedestrian accidents in the early morning hours.
-  - Other significant hours include `hour2 × pedestrian` (29.594) and `hour5 × pedestrian` (23.943), highlighting peak pedestrian accident hours.
+  - `hour1 × pedestrian`: 15.758, significant pedestrian accidents in the early morning hours.
+  - Other significant hours include `hour2 × pedestrian` (29.368) and `hour5 × pedestrian` (22.835), highlighting peak pedestrian accident hours.
 

### 2024-12-04 paper/paper.qmd
Added subtitle

@@
 ---
-title: "Predicting Fatalities in Toronto Car Crashes: Analyzing the Impact of Time, Vehicle Type, and Other Collision Determinants"
+title: "Predicting Fatalities in Toronto Car Crashes: Analyzing the Impact of Time, Vehicle Type, and Other Collision Determinants from 2015 to 2024"
+subtitle: "Vehicle Accidents Involving Motorcyclists and Pedestrians in the Evening Significantly Increases Fatality Risk"
 author: 

### 2024-12-10 paper/paper.qmd
spell check

@@
 For example, in 2024, thirty people have died on Toronto's roadways so far, which is a 20% increase from last year [@cbc_2024].
-However, news headlines, such as CBCs, highlight that Motor Vehicle Collisions (MVC) tend to either be generalized as a summary statistic or typically fatal events are over analyzed to the extent that environmental elements surrounding the crash are ignored.
+However, news headlines, such as CBC's, highlight that Motor Vehicle Collisions (MVC) tend to either be generalized as a summary statistic or typically fatal events are over analyzed to the extent that environmental elements surrounding the crash are ignored.
 General environmental elements such as the time of the crash or if a bicycle was involved in the accident are important to understand what increases the likelihood of a fatality occurring in an MVC.
@@
 The paper plans to include the police division variable as a general proxy for location and account if certain areas are more susceptible to crashes than others.
-@fig-division and @tbl-division shows the distirbution and breakdown of MVCs among police departments.Based on the @tbl-division, there seems to be no discernible pattern but D41 and D42 have the highest number of vehicular fatalities at 63 and 66, respectively.
+@fig-division and @tbl-division shows the distribution and breakdown of MVCs among police departments.Based on the @tbl-division, there seems to be no discernible pattern but D41 and D42 have the highest number of vehicular fatalities at 63 and 66, respectively.
 
@@
 #| label: tbl-division
-#| tbl-cap: Summary of the Police Divison Variable 
+#| tbl-cap: Summary of the Police Division Variable 
 #| tbl-align: center
@@
 To enhance the model, Bayesian priors were applied, introducing regularization and incorporating plausible ranges grounded.
-For the coefficient priors $\beta$, a normal distribution with a mean of 0 and a scale of 2.5 (autoscaled) was chosen to provide flexibility while mitigating overfitting.
+For the coefficient priors $\beta$, a normal distribution with a mean of 0 and a scale of 2.5 (autoscaled) was chosen to provide flexibility while mitigating over-fitting.
 Similarly, the intercept uses a normal prior with a mean of 0 and a scale of 2.5 to stabilize model estimates.
@@
 Fatalities are rare events, and the randomness inherent in accident outcomes—such as variations in road conditions, driver behavior, or other small variables—makes it difficult for models to predict them with high precision.
-Additionally, biases in data collection, like inconsistent reporting of certain variables or underreporting of less severe accidents, may lead to seleciton bias towards more severe accidents in our datasets, affecting our predictions.
+Additionally, biases in data collection, like inconsistent reporting of certain variables or under-reporting of less severe accidents, may lead to selection bias towards more severe accidents in our datasets, affecting our predictions.
 For example, in the United States, almost a third of car crashes involving a vulnerable road user go unreported [@streetsblog2021].
@@
 Fail to Remain Collisions occur when an individual involved in a MVC fails to stop and provide their information at the scene of a collision.
-Property Damage Collisions occur when an individual’s property has been damaged in a MVC or the value of damages is less than $2,000 for all involved These defintions, specifically the one for Fatal Collisions, sets a clear boundary for data inclusion, but excludes motor vehicle accidents that occur on private property.
+Property Damage Collisions occur when an individual’s property has been damaged in a MVC or the value of damages is less than $2,000 for all involved These definitions, specifically the one for Fatal Collisions, sets a clear boundary for data inclusion, but excludes motor vehicle accidents that occur on private property.
 
@@
 A self-reporting mechanism, such as an online portal or app, would allow drivers and witnesses to log details of these incidents conveniently.
-Additionally, redefining fatality criteria to include deaths occurring beyond 30 days due to collision-related injuries would align the dataset with public health standards, ensuring fatalities are not undercounted.
+Additionally, redefining fatality criteria to include deaths occurring beyond 30 days due to collision-related injuries would align the dataset with public health standards, ensuring fatalities are not under-counted.
 
@@
 Real-time telemetry from vehicles, collected via partnerships with automakers and insurers, can provide precise information on collision dynamics such as speed, braking patterns, and collision forces.
-Crowdsourcing data through a city-wide mobile app could also capture unreported incidents and near-misses, enriching the dataset with granular environmental and behavioral details.
+Crowd-sourcing data through a city-wide mobile app could also capture unreported incidents and near-misses, enriching the dataset with granular environmental and behavioral details.
 Integration of traffic camera data, analyzed using AI-powered algorithms, would add an unbiased and continuous data stream to validate collision reports and identify patterns.

### 2024-12-13 paper/paper.qmd
Updated Captions for Readability

@@
 #| label: fig-fatalities-cleaned
-#| fig-cap: Distribution of Fatalities in the Cleaned Dataset   
+#| fig-cap: Distribution of Fatalities Variable in the Cleaned Dataset    
 #| echo: false
@@
 #| label: tbl-hour
-#| tbl-cap: Summary of the Hour Variable 
+#| tbl-cap: Number of MVCs for Each Hour  
 #| tbl-align: center
@@
 #| label: fig-hour
-#| fig-cap: Distribution of the Hour Variable  
+#| fig-cap: Graphical Representation of Number of Fatalities for Each Hour 
 #| echo: false
@@
 
-### Type of Crash {#sec-data-predictor-crash}
+### Collision Type (Injury, Fail to Remain, and Property Damage Collisions) {#sec-data-predictor-crash}
 
@@
 #| label: fig-crashtype
-#| fig-cap: Breakdown of Each Collision Type by Fatalities  
+#| fig-cap: Breakdown of Distribution of Each Collision Type  
 #| echo: false
@@
 #| label: tbl-crashtype
-#| tbl-cap: Breakdown of MVCs Into the Different Categories, Broken Down by Fatalities
+#| tbl-cap: Breakdown of MVCs Into the Different Crash Type Categories by Fatalities
 #| tbl-align: center
@@
 ```
-
 ### Automobile {#sec-data-predictor-automobile}
@@
 #| label: fig-automobile
-#| fig-cap: Distribution of the Automobile Variable  
+#| fig-cap: Graphical Representation of the Automobile Variable's Values 
 #| echo: false
@@
 #| label: tbl-automobile
-#| tbl-cap: Summary of the Automobile Variable 
+#| tbl-cap: Breakdown of Automobile Variable Based on Fatalities Occured
 #| tbl-align: center
@@
 #| label: fig-motorcycle
-#| fig-cap: Distribution of the Automobile Variable  
+#| fig-cap: Graphical Representation of the Motorcycle Variable's Values
 #| echo: false
@@
 
-# Plot distribution of Hour Variable
+# Plot distribution of Motorcycle Variable
 ggplot(analysis_data, aes(x = motorcycle)) +
@@
 #| label: tbl-motorcycle
-#| tbl-cap: Summary of the Automobile Variable 
+#| tbl-cap: Summary of the Motorcycle Variable Based on Fatality Occurrence  
 #| tbl-align: center
@@
 #| label: fig-passenger
-#| fig-cap: Distribution of the Passenger Variable  
+#| fig-cap: Graphical Representation of the Passenger Variable's Values
 #| echo: false
@@
 
-# Plot distribution of Hour Variable
+# Plot distribution of Passenger Variable
 ggplot(analysis_data, aes(x = passenger)) +
@@
 #| label: tbl-passenger
-#| tbl-cap: Summary of the Passenger Variable 
+#| tbl-cap: Summary of the Passenger Variable Based on Fatality Occurrence 
 #| tbl-align: center
@@
 
-# Calculate summary statistics for the 'hour' variable
+# Calculate summary statistics for the 'passenger' variable
 summary_stats <- analysis_data %>%
@@
 #| label: tbl-pedestrian
-#| tbl-cap: Summary of the Pedestrian Variable 
+#| tbl-cap: Summary of the Pedestrian Variable Based on Fatality Occurrence
 #| tbl-align: center
@@
 
-# Calculate summary statistics for the 'hour' variable
+# Calculate summary statistics for the 'pedestrian' variable
 summary_stats <- analysis_data %>%
@@
 #| label: fig-pedestrian
-#| fig-cap: Distribution of the Pedestrian Variable  
+#| fig-cap: Graphical Representation of the Pedestrian Variable's Values
 #| echo: false
@@
 
-# Plot distribution of Hour Variable
+# Plot distribution of Pedestrian Variable
 ggplot(analysis_data, aes(x = pedestrian)) +
@@
 #| label: fig-division
-#| fig-cap: Distribution of the Police Division
+#| fig-cap: Number of Collisions per Police Division 
 #| echo: false
@@
 
-# Plot distribution of Hour Variable
+# Plot distribution of Police Division Variable
 ggplot(analysis_data, aes(x = police_division)) +
@@
 #| label: tbl-division
-#| tbl-cap: Summary of the Police Division Variable 
+#| tbl-cap: Summary of the Police Division Variable by Fatality Occurrence
 #| tbl-align: center
@@
 #| label: fig-hour-motorcycle
-#| fig-cap: Number of Fatalities by Hour for Motorcyclists
+#| fig-cap: Graphical Representation of Fatalities by Hour for Motorcyclists
 #| echo: false
@@
 #| label: tbl-hour-pedestrian
-#| tbl-cap: Summary of Fatalities by Hour for Motorcyclists and Pedestrians
+#| tbl-cap: Graphical Representation of Fatalities by Hour for Pedestrians
 #| tbl-align: center

### 2024-12-14 paper/paper.qmd
small paper edits for clarity

@@
 
-The variability in the coefficients for predictors l suggests that there may be additional, unobserved variables influencing the likelihood of fatalities that are not captured by the current model.
+The variability in the coefficients for predictors suggests that there may be additional, unobserved variables influencing the likelihood of fatalities that are not captured by the current model.
 This uncertainty could stem from gaps in the available data or from the limitations in the types of variables collected, such as the lack of detailed road conditions, driver behavior, or vehicle-specific characteristics (e.g., model, age of the vehicle).
